### 2017-06-02 KellyDouglass/Thesis/Tables/Paper1/XMPtable.tex

\begin{table}
\centering

\begin{tabu} to 0.75\linewidth {ccccccc}
Index\footnote[1]{KIAS-VAGC galaxy index number} & R.A. & Decl. & Redshift & \multicolumn{2}{c}{$12 + \log \left( \frac{\text{O}}{\text{H}} \right)$} & Void/Wall \\
\hline \\
268470 & \RA{13}{18}{17}{.82} & +\dec{02}{12}{59}{.83} & 0.0252 & 7.06 & $\pm$0.37 & Void \\
1422637 & \RA{14}{18}{12}{.14} & +\dec{13}{59}{33}{.98} & 0.0261 & 7.15 & $\pm$0.41 & Wall \\
839665 & \RA{08}{09}{53}{.53} & +\dec{29}{17}{04}{.82} & 0.0281 & 7.18 & $\pm$0.44 & Void \\
1168448 & \RA{11}{06}{41}{.00} & +\dec{45}{19}{09}{.28} & 0.0220 & 7.19 & $\pm$0.46 & Wall \\
1299291 & \RA{12}{17}{14}{.02} & +\dec{43}{18}{53}{.36} & 0.0233 & 7.21 & $\pm$0.42 & Wall \\
1170573 & \RA{11}{05}{39}{.42} & +\dec{46}{03}{28}{.37} & 0.0250 & 7.24 & $\pm$0.34 & Wall \\
2288717 & \RA{10}{46}{12}{.18} & +\dec{21}{31}{37}{.37} & 0.0248 & 7.27 & $\pm$0.48 & Wall \\
955643 & \RA{11}{42}{03}{.02} & +\dec{49}{21}{25}{.18} & 0.0244 & 7.28 & $\pm$0.44 & Wall \\
1344311 & \RA{12}{33}{13}{.64} & +\dec{11}{10}{28}{.46} & 0.0245 & 7.29 & $\pm$0.50 & Wall \\
1254352 & \RA{13}{29}{02}{.45} & +\dec{10}{54}{55}{.80} & 0.0237 & 7.30 & $\pm$0.44 & Wall \\
1857820 & \RA{08}{45}{00}{.34} & +\dec{27}{16}{47}{.04} & 0.0257 & 7.31 & $\pm$0.48 & Wall \\
866876 & \RA{09}{04}{57}{.96} & +\dec{41}{29}{36}{.42} & 0.0240 & 7.32 & $\pm$0.40 & Wall \\
833588 & \RA{08}{43}{10}{.71} & +\dec{43}{08}{53}{.58} & 0.0245 & 7.34 & $\pm$0.41 & Wall \\
283263 & \RA{14}{14}{12}{.88} & +\dec{01}{50}{12}{.88} & 0.0255 & 7.35 & $\pm$0.43 & Wall \\
184308 & \RA{09}{39}{09}{.38} & +\dec{00}{59}{04}{.15} & 0.0244 & 7.36 & $\pm$0.43 & Wall \\
1389829 & \RA{14}{31}{01}{.38} & +\dec{38}{04}{21}{.50} & 0.0269 & 7.41 & $\pm$0.46 & Wall \\
858951 & \RA{09}{31}{39}{.60} & +\dec{49}{49}{56}{.85} & 0.0251 & 7.46 & $\pm$0.46 & Wall \\
1270221 & \RA{13}{27}{39}{.85} & +\dec{50}{54}{09}{.69} & 0.0295 & 7.49 & $\pm$0.43 & Wall \\
431383 & \RA{08}{58}{44}{.96} & +\dec{50}{29}{58}{.98} & 0.0230 & 7.52 & $\pm$0.60 & Void \\
75442 & \RA{13}{13}{24}{.25} & +\dec{00}{15}{02}{.95} & 0.0264 & 7.55 & $\pm$0.35 & Void \\
1322765 & \RA{14}{15}{05}{.58} & +\dec{36}{22}{57}{.77} & 0.0273 & 7.57 & $\pm$0.40 & Wall\\
\end{tabu}

\caption[Extreme low-metallicity dwarf galaxies]{Details of the 21 extreme low gas-phase metallicity ($12 + \log(\text{O}/\text{H}) < 7.6$) galaxies found.  Four of these galaxies are located in voids (about 10\% of the void dwarf population measured) and seventeen are in more dense regions (about 19\% of the wall dwarf population measured); thus, there does not seem to be a special population of extreme low-metallicity galaxies in voids.  Further study of these galaxies is recommended to confirm metallicity values and identify any shared characteristics.}

\label{tab:lowZ_P1}

\end{table}