### 2017-06-02 KellyDouglass/Thesis/Thesis_checklist.tex

\documentclass{article}

\begin{document}

% Title
\title{Thesis checklist}
\author{Kelly Douglass, Michael S. Vogeley, advisor}
\date{21 June 2016}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Checkist

\section*{Papers}
\begin{itemize}
    \item April 2016 --- Submitted first paper to ApJ
    \item June 2016 --- Resubmit Paper 1 to ApJ (after applying referee suggestions)
    \item July 2016 --- Submit Paper 2 --- environmental dependence on N/O abundance in dwarf galaxies
    \item September 2016 --- Submit Paper 3 --- metallicity trend paper
    \item Winter/Spring 2017 --- (possible) Paper 4 --- environmental dependence on abundances in SHELS galaxies
    \item Spring 2017 --- (possible) Paper 5 --- small-scale environmental dependence on abundances in dwarf galaxies
\end{itemize}

\section*{Research}
\begin{itemize}
    \item Collaboration with Margaret Geller, Jabran Zahid using Hectospec data of SHELS galaxies
    \item Study metallicity relation between dwarf galaxies and their nearest bright neighbor
    \item Look at star formation history (metallicity), star formation rates, neutral gas content, morphology as related to large-scale environment
    \item (possibly) Compare results to dark matter halo simulations
\end{itemize}

\end{document}