### 2017-06-02 KellyDouglass/Thesis/Thesis_sections/Paper2.tex

\chapter[N/O ratio in dwarf galaxies]{Large-scale environmental dependence of the abundance ratio of nitrogen to oxygen in blue, star-forming galaxies fainter than $L_*$}

This paper is published in the \emph{Astrophysical Journal}, 2017 (Vol. 837, 
pages 42--55).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    INTRODUCTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% Production of N in stars (primary v. secondary) and its relation to O
The measurement of the abundance of heavier elements relative to hydrogen in a 
galaxy can indicate the galaxy's evolutionary stage.  As stars evolve, they 
slowly convert hydrogen into heavier elements, increasing the ratio of the 
heavier elements (oxygen, nitrogen, etc.) to hydrogen.  The ratio of oxygen to 
hydrogen is often used to determine the chemical evolution of a galaxy because 
oxygen is the most abundant element in the universe (after hydrogen and helium) 
and because oxygen has very strong emission lines in the optical regime that 
cover a range of ionization states \citep{Kewley02}.

It is instructive to also study the relative abundances of the heavy elements in 
a galaxy.  Rather than indicating the amount of hydrogen converted to heavier 
elements, the ratio of two heavy elements can reveal important details about the 
nucleosynthesis process and the chemical conditions of the galaxy when the last 
star formation episode occurred \citep{Izotov99}.  One of the easiest and most 
informative ratios to study is nitrogen to oxygen.

From what we currently understand of stellar nucleosynthesis, we can group its 
products into two classes: primary and secondary elements.  The yields of 
primary elements (carbon and oxygen, for example) are independent of the initial 
metallicity of the star, while the yields of secondary elements depend on the 
initial abundance of heavy elements in the star.  Nitrogen is unique --- it can 
behave as both a primary and secondary element \citep{Matteucci86}.  Nitrogen is 
produced during the CNO cycle, which is one of the two main processes of 
hydrogen burning in a star.  The CNO cycle fuses four protons into a helium atom 
with two positrons and two electron neutrinos as by-products.  It tends to occur 
in more massive stars than our Sun, due to the higher temperature required for 
the fusion processes involved.  Carbon is a catalyst of the CNO cycle, not a 
product.  As a result, if carbon is not initially present within the star, then 
nitrogen is produced in the same relative abundance as carbon and oxygen --- 
nitrogen behaves as a primary element.  However, if the interstellar medium 
(ISM) has a relatively high abundance of heavier elements from previous star 
formation episodes, then nitrogen behaves as a secondary element, since its 
production is based on carbon and oxygen produced prior to the star's creation.

The majority of the production of oxygen and nitrogen is thought to occur in 
different mass stars --- nitrogen is produced in the CNO cycle of 
intermediate-mass stars ($4M_{\odot} < M_* < 8M_{\odot}$), while oxygen is 
primarily produced in the helium-, carbon-, and neon-burning stages of 
higher-mass stars ($M_* > 4M_{\odot}$) \citep{Henry00,Henry06}.  The CNO cycle 
can occur in lower-mass stars (the minimum temperature is only 
$1.5\times 10^7$K), but it requires carbon as a catalyst.  If there is already 
carbon present in a star at its birth, the CNO cycle can commence much earlier 
in the star's lifetime than if it is composed primarily of hydrogen at its 
birth.

% Why are we looking at N/O?  What does a high v. low N/O mean?
A measurement of the N/O ratio indicates where a galaxy is in its chemical 
evolution.  The relative amounts of these two elements can be influenced by 
nucleosynthesis, a galaxy's star formation history, and/or a varying initial 
mass function (IMF), for example.  The star formation history of a galaxy can be 
strongly influenced by the galaxy's environment.  Galactic interactions can 
cause bursts of star formation in addition to secular star formation.  Due to 
the time delay in the release of nitrogen and oxygen from the stellar 
population, galaxies that have more recently experienced star formation will 
result in lower N/O ratios (since oxygen is released sooner than nitrogen, due 
to higher-mass stars being responsible for the production of oxygen).  In 
addition to this time delay, if a galaxy has enough heavy elements present in 
its gas at the time of the stars' births, secondary nitrogen will be produced in 
addition to primary.  This would result in higher N/O ratios, and there would be 
a correlation between the metallicity and the N/O ratio in the galaxies.

% Large-scale environment overview (voids, how they were discovered)
Large galaxy redshift surveys have shown that the large-scale structure of 
galaxies is similar to that of a three-dimensional cosmic web \citep{Bond96}, 
where voids (large, underdense regions that occupy approximately 60\% of space) 
separate galaxy clusters that are connected by thin filaments of galaxies.  
% Why we think the environment should influence N/O, especially in dwarf galaxies
These cosmic voids are an important environment for studying galaxy formation 
\citep[see][for a review]{vandeWeygaert11}, as the $\Lambda$CDM cosmology 
predicts void galaxies to have lower mass and be retarded in their star 
formation when compared to those in denser environments 
\citep[e.g.,][]{Gottlober03, Goldberg05, Cen11}.  Because dwarf galaxies are 
sensitive to many astrophysical effects, including cosmological reionization, 
internal feedback from supernovae and photoheating from star formation, external 
effects from tidal interactions and ram pressure stripping, small-scale details 
of dark matter halo assembly, and properties of dark matter, they should be the 
most sensitive to the effects of the void environment.

Previous work by \cite{Douglass17a} (hereafter Paper I) shows that there is no 
large-scale environmental dependence of the amount of oxygen in dwarf galaxies, 
in contrast to earlier studies by \cite{Pustilnik06}, \cite{Cooper08}, 
\cite{Deng11}, and \cite{Filho15}, for example.  One of the main arguments for 
the existence of an environmental dependence of the metallicity of galaxies 
centers around the idea that void galaxies are surrounded by pristine hydrogen 
that is unavailable to galaxies in denser regions.  By looking at just N/O, we 
remove the hydrogen dependence of the relative abundances.  Detecting a 
difference in the N/O ratio due to the large-scale environment would indicate 
that the cosmic environment has some influence on the nucleosynthesis of 
secondary elements.  In addition, if the environment does have some very minor 
effect on the metallicity of a galaxy, removing the hydrogen dependence could 
amplify this effect above the noise of the data.  Combined with the metallicity 
results in Paper I, we might be able to discern a large-scale environmental 
effect on the chemical evolution of galaxies.

% Why this paper is significant, new, and important (large data set - statistically relevant)
Large-scale sky surveys like the Sloan Digital Sky Survey 
\citep[SDSS;][]{Abazajian09} contain a large sample of dwarf galaxies, allowing 
us to analyze the dwarf galaxy population in the nearby universe with more 
statistical significance.  Over 1000 voids have been identified in SDSS DR7 
\citep{Pan12}, and SDSS provides spectroscopy to permit abundance estimates of 
those dwarf galaxies found in these voids.  Thus, we are able to estimate the 
N/O ratio as a function of large-scale environment for the largest sample of 
dwarf galaxies to date.

% Why are we using MPA-JHU emission line measurements?
We make use of the MPA-JHU catalog's reprocessed spectroscopic 
data\footnote{Available at http://www.mpa-garching.mpg.de/SDSS/DR7/} to study 
the N/O abundance ratio of a large collection of dwarf galaxies in SDSS DR7.  
Because our analysis depends on the weak [O{\sc iii}] $\lambda 4363$ auroral 
line, the MPA-JHU catalog's more detailed treatment of the stellar continuum 
permits the weaker emission lines to become more apparent.  As a result, using 
this catalog's flux measurements should improve the accuracy of our results.  We 
study the N/O abundance ratio of these dwarf galaxies as a function of 
large-scale environment to discern whether the large-scale environment has an 
effect on the relative abundance of heavier elements in dwarf galaxies.

% Outline of paper
Our paper is organized as follows.  Section 2 describes the method used to 
estimate the chemical abundances in galaxies.  We remind the reader of the 
source of our data in Section 3.  Section 4 includes the results of our analysis, 
and Section 5 is a discussion of the implications of our results on the 
large-scale environmental effects on galaxy evolution.  Finally, Section 6 
summarizes our conclusions and discusses future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    THEORY
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Theory]{Estimation of gas-phase chemical abundances from optical spectroscopy}
\label{sec:Theory}

We study a galaxy's oxygen and nitrogen abundances because they are relatively 
abundant elements, they emit strong lines in the optical regime (including for 
several ionization states in oxygen), and a ratio of some of the oxygen lines 
provides a good estimate of the electron temperature \citep{Kewley02}.  What 
follows is a description of the theory and methods we employ to estimate the 
oxygen and nitrogen abundances in dwarf galaxies.

% ------------------------------------------------------------------------------
\subsection{[N{\sc ii}]}

The energy-level diagram for the various transitions of [N{\sc ii}] is very 
similar to that of [O{\sc iii}], since they have the same electron ground-state 
configuration ((1s)$^2$(2s)$^2$(2p)$^2$).  The similarities can be seen in Fig. 
\ref{fig:transitions_P2}.  Therefore, an estimate of the electron temperature can 
be made from the [N{\sc ii}] $\lambda 5755$ emission line.  However, this line 
is weaker than the [O{\sc iii}] $\lambda 4363$ auroral line (since there is less 
N than O in galaxies), so we use the [O{\sc iii}] auroral line for our 
temperature estimates, as in Paper I.  After obtaining a temperature and density 
estimate, we use the [N{\sc ii}] $\lambda \lambda 6548, 6584$ doublet to 
estimate the abundance of singly ionized nitrogen in a galaxy.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper2/NIIOIII_energy_level_diagram-figure0}
    \caption[{[\ion{O}{3}] and [\ion{N}{2}] energy-level diagram}]{Energy-level 
    diagram for [O{\sc iii}] and [N{\sc ii}] $(2p^2)$ ions.  The most important 
    transitions are shown; all are in the visible spectrum.  These forbidden 
    transitions in both oxygen and nitrogen provide an estimate of the electron 
    temperature in the interstellar gas.  Because oxygen is more abundant, we 
    use the oxygen lines to estimate the temperature of the gas.}
    \label{fig:transitions_P2}
\end{figure}

% ------------------------------------------------------------------------------
\subsection{Direct $T_e$ method}

We use the same method to calculate the nitrogen abundance as in Paper I to 
estimate the oxygen abundance.  However, here we use the [N{\sc ii}] 
$\lambda \lambda 6548, 6584$ doublet instead of the [O{\sc ii}] $\lambda 3727$ 
and [O{\sc iii}] $\lambda \lambda 4959,5007$ doublets.  Because the temperature 
estimate depends on the auroral line [O{\sc iii}] $\lambda 4363$, this method is 
often difficult to employ.  As a result, it works best with low-redshift, 
low-metallicity galaxies.  The electron temperature is derived by solving the 
following system of equations:
\begin{equation}
	t_3 = \frac{1.432}{\log[(\lambda 4959 + \lambda 5007)/\lambda 4363] - \log C_T}
\end{equation}
where $t_3 = 10^{-4} T_e(\text{O}^{++})$ and
\begin{equation}
	C_T = (8.44 - 1.09t_3 + 0.5t_3^2 - 0.08t_3^3)\frac{1 + 0.0004x_3}{1 + 0.044x_3}
\end{equation}
where $x_3 = 10^{-4} n_e t_3^{-0.5}$.  The ionic abundances are then found with 
the equations
\begin{align}
    12 + \log \left( \frac{\text{O}^{++}}{\text{H}^+} \right) &= \log \frac{\lambda 4959 + \lambda 5007}{\text{H}\beta} + 6.200 + \frac{1.251}{t_3} - 0.55\log t_3 - 0.014t_3 \\
	12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) &= \log \frac{\lambda 3727}{\text{H}\beta} + 5.961 + \frac{1.676}{t_2} - 0.40\log t_2 - 0.034t_2 + \log (1+1.35x_2) \\
	12 + \log \left( \frac{\text{N}^+}{\text{H}^+} \right) &= \log \frac{\lambda 6548 + \lambda 6584}{\text{H}\beta} + 6.234 + \frac{0.950}{t_2} - 0.42\log t_2 - 0.027t_2 + \log (1 + 0.116x_2)
\end{align}
where $t_2 = 10^{-4} T_e(\text{O}^+)$ and $x_2 = 10^{-4} n_e t_2^{-0.5}$.  We 
assume that $T_e(\text{N}^+) = T_e(\text{O}^+)$.

The signal-to-noise ratio of the SDSS spectra is too low to directly estimate 
the temperature of the gas in the low-ionization zone.  As a result, we use the 
relation $t_2 = 0.7t_3 + 0.3$ by \cite{Garnett92}.  This relation has been shown 
to overestimate this temperature \citep{Andrews13}.  Since the metal emission 
lines are the primary method of cooling for the gas, a high temperature 
corresponds to a low metallicity.  Therefore, an overestimate of the temperature 
results in an underestimated abundance.  As shown in Paper I, this only affects 
perhaps 15 of the dwarf galaxies in our sample and does not influence our 
conclusions.

The sum of the abundances of each of the element's ionization states is equal to 
the total abundance of any element, whether or not all ionization states are 
observed.  Most of oxygen exists as either singly or doubly ionized, so the 
total oxygen abundance is
\begin{equation}
	\frac{\text{O}}{\text{H}} = \frac{\text{O}^{++}}{\text{H}^+} + \frac{\text{O}^+}{\text{H}^+}
\end{equation}
Since we can only observe the nitrogen abundance in one of the main ionization 
states, we use an ionization correction factor (ICF) to account for the missing 
states.  For any element X, the total abundance is 
\begin{equation}\label{eqn:ICF}
    \frac{\text{X}}{\text{H}} = \sum_i ICF_i \frac{\text{X}^i}{\text{H}}
\end{equation}
For nitrogen, we employ the ICFs as defined in \cite{Izotov06}:
\begin{equation}\label{eqn:ICF_N}
    ICF(\text{N}^+) = \left\{ \begin{array}{ll}
    -0.825v + 0.718 + \frac{0.853}{v} & \mbox{low $Z$}\\
    -0.809v + 0.712 + \frac{0.852}{v} & \mbox{intermed $Z$}\\
    1.467v + 1.752 + \frac{0.688}{v} & \mbox{high $Z$}
    \end{array} \right.
\end{equation}
where $v = \text{O}^+ / (\text{O}^+ + \text{O}^{++})$.  The range for low $Z$ 
covers galaxies with $12 + \log \left( \text{O}/\text{H} \right) \leq 7.2$, 
while high $Z$ includes galaxies with 
$12 + \log \left( \text{O}/\text{H} \right) \geq 8.2$.  For galaxies with 
$7.2 < 12 + \log \left( \text{O}/\text{H} \right) < 7.6$, the values for the 
ICFs are a linear interpolation between the low-$Z$ and intermediate-$Z$ values, 
while the ICFs for galaxies with 
$7.6 < 12 + \log \left( \text{O}/\text{H} \right) < 8.2$ are a linear 
interpolation between the intermediate-$Z$ and high-$Z$ values.

The N/O ratio can be found from the O/H and N/H ratios:
\begin{equation}\label{eqn:NO}
    \log \left(\frac{\text{N}}{\text{O}}\right) = \left[ 12 + \log \left( \frac{\text{N}}{\text{H}} \right) \right] - \left[ 12 + \log \left( \frac{\text{O}}{\text{H}} \right) \right]
\end{equation}

%The N$^+$/O$^+$ ratio can then be found from the O$^+$/H and N$^+$/H ratios:
%\begin{equation}
%    \log \left(\frac{\text{N}^+}{\text{O}^+}\right) = \left[ 12 + \log \left( \frac{\text{N}^+}{\text{H}^+} \right) \right] - \left[ 12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) \right]
%\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    DATA
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Data]{SDSS data and galaxy selection}

The SDSS Data Release 7 \citep[DR7;][]{Abazajian09} uses drift scanning to map 
approximately one-quarter of the northern sky; it is a wide-field multiband 
imaging and spectroscopic survey.  A dedicated 2.5 m telescope at the Apache 
Point Observatory in New Mexico \citep{Fukugita96, Gunn98} takes the photometric 
data in the five-band SDSS system --- $u$, $g$, $r$, $i$, and $z$.  Galaxies 
selected for spectroscopic analysis must have a Petrosian $r$-band magnitude 
$m_r < 17.77$ \citep{Lupton01, Strauss02}.  Two double fiber-fed spectrographs 
and fiber plug plates take the spectra in an observed wavelength range of 
3800--9200 \AA with a resolution $\lambda / \Delta \lambda \sim 1800$ and a 
minimum fiber separation of 55" \citep{Blanton03}.  As in Paper I, we use 
the emission-line flux data from the MPA-JHU value-added catalog, which is based 
on the SDSS DR7 sample of galaxies.  Total star formation rates and total 
specific star formation rates are also from the MPA-JHU value-added catalog, 
following the technique discussed in \cite{Brinchmann04}.  The MPA-JHU catalog 
is also the source of the stellar mass estimates used, as calculated in 
\cite{Tremonti04}, following the method outlined in \cite{Kauffmann03}.  The 
KIAS value-added galaxy catalog \citep{Choi10} is our source of the absolute 
magnitudes and colors of the galaxies.

%-------------------------------------------------------------------------------
\subsection{Spectroscopic selection}\label{sec:SDSS_limits}

The following requirements are implemented on the SDSS DR7 main spectroscopic 
galaxy sample described above.  We use the same requirements for our sample as 
in Paper I; all galaxies must have
\begin{enumerate}
    \item{$M_r > -17$ (dwarf galaxies);}
    \item{a minimum $5\sigma$ detection of H$\beta$;}
    \item{a minimum $1\sigma$ detection of [\ion{O}{3}] $\lambda 4363$;}
    \item{a flux $>0$ for all other required lines;}
    \item{$T_e (\text{\ion{O}{3}}) < 3\times 10^4 \text{ K}$;}
    \item{a star-forming BPT classification by \cite{Brinchmann04}.}
\end{enumerate}
We also use the \texttt{oii\_flux} value from the MPA-JHU catalog in place of 
their [\ion{O}{2}] $\lambda \lambda 3726, 3729$ flux measurement since we are 
working at such low redshifts ($0.02 < z < 0.03$).  Detailed descriptions of 
these criteria can be found in Section 3.1 of Paper I.

%-------------------------------------------------------------------------------
\subsection{Void classification}

The large-scale environment of the galaxies was determined using the void 
catalog constructed by \cite{Pan12}, which is based on the galaxies in the SDSS 
DR7 catalog.  The VoidFinder algorithm of \cite{Hoyle02} removes all isolated 
galaxies with absolute magnitudes $M_r < -20$ (a galaxy is defined to be 
isolated if its third nearest neighbor is more than 7 $h^{-1}$ Mpc away).  
Placing a grid over the remaining galaxies, VoidFinder grows spheres in the 
centers of all grid cells that contain no galaxies.  The spheres expand until 
they encounter four galaxies on the surface.  To be considered part of a void, a 
sphere must have a minimum radius of 10 Mpc; two spheres that overlap by more 
than 10\% are considered part of the same void.  We refer the reader to 
\cite{Hoyle02} for a more detailed description of the VoidFinder algorithm.  
Using these voids, galaxies that live within any void sphere are classified as 
a void galaxy; those that are outside the spheres are considered wall galaxies.  
Due to the construction of the void spheres, we cannot identify any voids within 
10 Mpc of the edge of the survey.  As a result, the large-scale environment of 
any galaxy within this boundary is uncertain.

9519 of the $\sim$800,000 galaxies with spectra available in SDSS DR7 are dwarf 
galaxies ($M_r > -17$).  42 void dwarf galaxies, 89 wall dwarf galaxies, and 4 
dwarf galaxies with uncertain large-scale environments are left to analyze after 
applying the spectroscopic cuts (or 135 dwarf galaxies in total, 131 of which 
are used in the environmental study).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    ANALYSIS & RESULTS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Results]{Abundance analysis and results}

Our primary objective is to perform a relative measurement of the N/O ratio of 
dwarf galaxies to discern how the large-scale environment affects their chemical 
evolution.  As discussed in Paper I, multiple methods have been developed for 
metallicity calculations based on the quality of the spectra.  We use only the 
direct $T_e$ method for our abundance calculations, due to the limited galaxy 
types used in the calibration or theoretical development of other methods.

For reference, the solar metallicity $Z_{\odot} = 8.69\pm 0.05$ 
\citep{Asplund09}.

%-------------------------------------------------------------------------------
\subsection{Estimation of uncertainties and comparison of N/O and N$^+$/O$^+$}

% Uncertainties
We estimate uncertainties in the computed abundances using a Monte Carlo method.  
We calculate 100,000 abundance estimates using the measured line fluxes and 
scaled uncertainty estimates.  A new positive ``fake'' line flux is drawn from a 
normal distribution for each abundance estimate.  The standard deviation in 
the sets of 100,000 calculated abundance values is used for the error in the 
abundance calculation.  A more in-depth description of this process can be found 
in Paper I.

It has been common practice to assume that N/O $\cong$ N$^+$/O$^+$, thus 
eliminating the need for the ICF in Eqn. \ref{eqn:ICF}.  We find that this is a 
reasonable but slightly biased approximation, agreeing with the results of 
\cite{Nava06}.  A comparison of the N/O ratio and the N$^+$/O$^+$ ratio for our 
set of dwarf galaxies can be seen in Fig. \ref{fig:NO_NpOp}; galaxies with 
absolute magnitudes $M_r > -20$ are shown in gray for context.  A linear fit to 
all star-forming galaxies with magnitudes $M_r > -20$ has a slope of only 
$0.927\pm 0.0018$, with an rms error of 0.032 for the fit.  This comparison 
indicates that lower values of the N$^+$/O$^+$ ratio underestimate the N/O 
ratio, while higher values of N$^+$/O$^+$ overestimate the N/O ratio.  
Throughout this paper, we study the N/O ratio using ICF-corrected estimates of 
N/O (Eqns. \ref{eqn:ICF_N} and \ref{eqn:NO}).

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper2/1sig_I06_dwarf_0-20_SF_t3_logNpOp_logNO_fit}
    \caption[N/O versus N$^+$/O$^+$]{Comparison of the N/O and N$^+$/O$^+$ 
    abundance ratios for our dwarf galaxies.  Brighter galaxies ($M_r > -20$) 
    are shown in gray for context.  All our star-forming galaxies with 
    $M_r > -20$ roughly follow the approximation that N/O $\cong$ N$^+$/O$^+$, 
    which is often assumed in other studies of the abundance ratio of nitrogen 
    to oxygen.  In this paper, we will be using the abundance ratio N/O for our 
    analysis.}
    \label{fig:NO_NpOp}
\end{figure}

%-------------------------------------------------------------------------------
\subsection{Sources of systematic error}

There is a radial dependence of many physical properties of galaxies 
\citep{Bell00}.  Consequently, abundance estimates may depend on the locations 
of the spectroscopic fiber on the galaxy.  If all of the galaxy's light is not 
contained within the fiber of the spectrograph, the estimated abundances will 
not necessarily be representative of global abundance values.  For example, 
\cite{Bell00} show that the metallicity is not constant throughout a galaxy.  
Due to the spatially resolved spectra produced by MaNGA of SDSS-IV 
\citep{SDSS13}, a statistically significant measure of the radial dependence of 
a galaxy's metallicity should soon be possible \citep{Wilkinson15}.  In SDSS 
DR7, the fiber diameter is 3", corresponding to a physical diameter between 1.29 
and 1.93 kpc at redshifts $0.02 < z < 0.03$.  This covers a majority of most 
dwarf galaxies' luminous surfaces.  The fiber is almost always placed on the 
brightest spot of the galaxy, which is often the center of the galaxy for spiral 
and ellipticals.  Since the metallicity has been shown to decrease at large 
radius, these abundance values may be overestimates of the global abundances.  
Since many dwarf galaxies are irregular galaxies, the fiber is instead focused 
on a bright \ion{H}{2} region.  As a result, we are estimating the abundances of 
the gas from which stars recently formed.

We are implicitly limiting our sample of galaxies to only blue, star-forming 
dwarf galaxies as a result of our selection criteria outlined in Section 
\ref{sec:SDSS_limits}.  Consequently, this is not a representative sample of the 
full dwarf galaxy population.  In this study we are only able to discuss the 
large-scale environmental influence on blue, star-forming dwarf galaxies within 
a narrow redshift range.  It is impossible to use the direct $T_e$ method to 
measure the chemical abundances of red dwarf galaxies because the UV photons 
from young stars are needed to excite the interstellar gas.

%-------------------------------------------------------------------------------
\subsection{Galaxy abundances}

Abundances estimated using the direct $T_e$ method for our dwarf galaxy sample 
are listed in Table \ref{tab:Results_P2}, along with other important 
characteristics and identification for the galaxies (including their large-scale 
environment classification).

% Results table
\input{Tables/Paper2/results_table_short}

%-------------------------------------------------------------------------------
\subsubsection{Oxygen and nitrogen abundances}\label{sec:OH_NH}

Histograms of the resulting oxygen and nitrogen abundances are shown in Figs. 
\ref{fig:met1sig} and \ref{fig:N_1sig}, respectively.  Both figures show very 
little difference in the distribution of abundance values in dwarf galaxies 
between voids and walls.  A two-sample Kolmogorov-Smirnov (K-S) test quantifies 
this observation --- it produced a test statistic of 0.13 for oxygen and 0.11 
for nitrogen, corresponding to a probability of 67.1\% and 83.8\%, respectively, 
that a test statistic greater than or equal to this calculated test statistic 
will be measured if the void sample were drawn from the wall sample.  The 
cumulative distribution function (CDF) of these samples can be seen in the right 
panel of Figures \ref{fig:met1sig} and \ref{fig:N_1sig}.  The K-S test 
quantifies the visual impression in these figures that the distributions of 
oxygen and nitrogen abundances are similar for dwarf galaxies in voids and 
walls.

The average and median values of the dwarf galaxy abundances indicate very 
little large-scale environmental influence on the oxygen and nitrogen 
abundances.  The average oxygen abundance for void dwarf galaxies is 
$7.99\pm 0.049$ and the median is 8.04, while the average for wall dwarf 
galaxies is $7.93\pm 0.036$ with a median value of 8.01.  This implies that the 
wall dwarf galaxies have lower oxygen abundances by an average of 
$0.07\pm 0.060$ relative to the void dwarf galaxies; the shift in the median 
values is 0.03 for the dwarf galaxies, with wall dwarf galaxies having lower 
oxygen abundances than void dwarf galaxies.  There is a also a shift in the 
nitrogen abundances for the dwarf galaxies: void dwarf galaxies have an average 
nitrogen abundance of $6.74\pm 0.035$ and a median of 6.77, while the wall dwarf 
galaxies have an average nitrogen abundance of $6.72\pm 0.025$ and a median of 
6.75.  Again, wall dwarf galaxies have, on average, $0.02\pm 0.043$ lower 
nitrogen abundances than the void dwarf galaxies (the median shift for the 
nitrogen abundance of dwarf galaxies is 0.01, with wall galaxies lower than void 
dwarf galaxies).  These shifts are within the uncertainty, so they are not 
statistically significant --- if there is a large-scale environmental influence 
on the abundances of oxygen and nitrogen relative to hydrogen in dwarf galaxies, 
it is small.

% Dwarf galaxies
%\begin{figure*}
%    \centering
%    \includegraphics[width=0.49\textwidth]{Images/Paper1/1sig_dwarf_SF_t3_12logOH_hist}
%    \includegraphics[width=0.49\textwidth]{Images/Paper1/1sig_dwarf_SF_t3_12logOH_CDF}
%    \caption{Gas-phase oxygen abundances (relative to hydrogen) of void dwarf 
%    (red solid line) and wall dwarf (black dashed line) galaxies, taken from 
%    Fig. 4 in Paper I.  As we demonstrate in Paper I, the gas-phase oxygen 
%    abundances in dwarf galaxies do not depend on the large-scale environment.}
%    \label{fig:met1sig}
%\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{Images/Paper2/1sig_dwarf_SF_t3_12logNH_hist}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/1sig_dwarf_SF_t3_12logNH_CDF}
    \caption[Nitrogen distribution of 135 dwarf galaxy sample]{Abundance of 
    nitrogen relative to hydrogen of void dwarf (red solid line) and wall dwarf 
    (black dashed line) galaxies.  A two-sample K-S test of the two data sets 
    results in an asymptotic $p$-value of 0.84, indicating an 84\% probability 
    that a test statistic greater than the observed value of 0.11 will be seen 
    if the void sample is drawn from the wall sample.  This is reflected 
    visually, as there appears to be very little difference between the two 
    populations, indicating that there is little large-scale environmental 
    influence on the nitrogen abundance of dwarf galaxies.}
    \label{fig:N_1sig}
\end{figure*}

%%%% Brighter star-forming galaxies %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To see how our results of the environmental dependence of dwarf galaxies compare 
with somewhat brighter galaxies, we perform the same analysis on galaxies with 
absolute magnitudes $-17 > M_r > -20$.  The results of this analysis can be seen 
in Figs. \ref{fig:Z_bright} and \ref{fig:N_bright}.  As the dwarf galaxies have 
already shown, there is no obvious large-scale environmental dependence of the 
oxygen and nitrogen abundances of these brighter galaxies.  The results of a 
two-sample K-S test (listed in Table \ref{tab:stats}) mostly support this 
conclusion.  In the brightest magnitude bin (galaxies with $-19 > M_r > -20$), 
the K-S test returns a $p$-value of only 0.00062 for the oxygen abundances, 
indicating only a 0.062\% chance that there will be a test statistic greater 
than 0.07 if the void sample is drawn from the wall sample.  The oxygen 
abundances for void galaxies are higher than the wall galaxies by an average of 
$0.04\pm 0.017$ in this absolute magnitude bin, reinforcing the results of the 
K-S test that there is a large-scale environmental influence on the oxygen 
abundance in galaxies with magnitudes $-19 > M_r > -20$.  While the results of 
the K-S test are not as convincing for this magnitude range in the nitrogen 
abundances, there is still an average shift of $0.02\pm 0.011$ toward higher 
nitrogen abundances for void galaxies.  While only one magnitude bin shows a 
statistically significant shift between the two environments, all magnitude bins 
are shifted in the same direction.  This trend suggests that there may be a mild 
influence on the chemical evolution of galaxies due to their large-scale 
environment.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Images/Paper2/1sig_17-20_SF_t3_12logOH_stacked}
    \caption[Metallicity distribution of star-forming galaxies with 
    $-17 > M_r > -20$]{Gas-phase oxygen abundances relative to hydrogen of void 
    (red solid line) and wall (black dashed line) star-forming galaxies with 
    $-17 > M_r > -18$ (top), $-18 > M_r > -19$ (middle), and $-19 > M_r > -20$ 
    (bottom).  The results of a two-sample K-S test of the two data sets in each 
    absolute magnitude range can be found in Table \ref{tab:stats}.  These 
    results are reflected visually, as there appears to be very little 
    difference between the two populations (regardless of absolute magnitude), 
    indicating that there is little large-scale environmental influence on the 
    oxygen abundance of star-forming galaxies with absolute magnitudes 
    $-17 > M_r > -20$.}
    \label{fig:Z_bright}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Images/Paper2/1sig_17-20_SF_t3_12logNH_stacked}
    \caption[Nitrogen distribution of star-forming galaxies with 
    $-17 > M_r > -20$]{Abundance of nitrogen relative to hydrogen of void (red 
    solid line) and wall (black dashed line) star-forming galaxies with 
    $-17 > M_r > -18$ (top), $-18 > M_r > -19$ (middle), and $-19 > M_r > -20$ 
    (bottom).  The results of a two-sample K-S test of the two data sets in each 
    absolute magnitude range can be found in Table \ref{tab:stats}.  These 
    results are reflected visually, as there appears to be very little 
    difference between the two populations (regardless of absolute magnitude), 
    indicating that there is little large-scale environmental influence on the 
    nitrogen abundance of star-forming galaxies with absolute magnitudes 
    $-17 > M_r > -20$.}
    \label{fig:N_bright}
\end{figure}

We note that there appears to be a shift in the oxygen and nitrogen abundances 
between the absolute magnitude bins in Figs. \ref{fig:Z_bright} and 
\ref{fig:N_bright} that is opposite to what is predicted by the mass-metallicity 
relation \citep{Tremonti04}.  This shift toward lower oxygen abundances as the 
galaxies increase in brightness is possibly due to the fact that the metallicity 
estimates are so dependent on the temperature-sensitive [\ion{O}{3}] 
$\lambda 4363$ auroral line.  As galaxies increase in metallicity, this line 
becomes weaker (as its strength is inversely proportional to the temperature).  
If the flux of this line is being underestimated, then the temperature is being 
overestimated, and therefore the oxygen and nitrogen abundances are being 
underestimated.  If we are seeing an underestimate of flux of the [\ion{O}{3}] 
$\lambda 4363$ emission line (and therefore an overestimate of the temperature 
in the region), then we should see a shift toward lower N/H values as the 
absolute magnitude is increased as well.  This pattern can be seen in Fig. 
\ref{fig:N_bright}.

%\begin{table}
%
%    \begin{tabu} in 0.75\textwidth {ccccccccc}
%        Absolute magnitude range & Environment & \# of galaxies & Average & Median & Average shift\footnote[1]{Wall -- Void (Positive shifts indicate that the wall galaxy average is greater than the void average; negative shifts indicate that the void galaxy average is greater than the wall average.)} & Median shift\footnote[1] & $p$-value & KS test statistic \\
%        \hline \\
%        \multirow{2}{*}{Dwarf galaxies} & Void & 42 & $6.74\pm 0.035$ & 6.77 & \multirow{2}{*}{$-0.02\pm 0.043$} & \multirow{2}{*}{0.8383} & \multirow{2}{*}{0.1129}\\
%         & Wall & 89 & $6.72\pm 0.025$ & 6.75 & & & \\
%        \multirow{2}{*}{$-17>M_r>-18$} & Void & 423 & $6.80\pm 0.011$ & 6.82 & \multirow{2}{*}{$0.00\pm 0.014$} & \multirow{2}{*}{0.7364} & \multirow{2}{*}{0.0401}\\
%         & Wall & 895 & $6.80\pm 0.008$ & 6.81 & & & \\
%        \multirow{2}{*}{$-18>M_r>-19$} & Void & 829 & $6.87\pm 0.009$ & 6.85 & \multirow{2}{*}{$0.01\pm 0.011$} & \multirow{2}{*}{0.0873} & \multirow{2}{*}{0.0538}\\
%         & Wall & 1498 & $6.89\pm 0.007$ & 6.87 & & & \\
%        \multirow{2}{*}{$-19>M_r>-20$} & Void & 798 & $6.97\pm 0.009$ & 6.94 & \multirow{2}{*}{$-0.02\pm 0.011$} & \multirow{2}{*}{0.2272} & \multirow{2}{*}{0.0455}\\
%         & Wall & 1490 & $6.95\pm 0.007$ & 6.93 & & & \\
%    \end{tabu}
%
%    \tablecomments{Statistics on the gas-phase nitrogen abundance in void and 
%    wall galaxies in each of the absolute magnitude ranges listed.  Combined 
%    with the histograms in Figs. \ref{fig:N_1sig} and \ref{fig:N_bright}, it is 
%    clear that there is no large-scale environmental dependence of the nitrogen 
%    abundance in galaxies.}
%
%    \label{tab:KStest_N}
%
%\end{table}

%-------------------------------------------------------------------------------
\subsubsection{Ratio of nitrogen to oxygen}

% Dwarf galaxies
In addition to studying the oxygen and nitrogen abundances relative to hydrogen, 
we also look at the ratio of nitrogen to oxygen.  The N/O abundance ratio 
suggests a slightly stronger environmental influence on the chemical evolution 
of dwarf galaxies than the oxygen and nitrogen abundances individually.  As can 
be seen in Fig. \ref{fig:NOratio}, there is a shift in the N/O ratio to lower 
values in the void dwarf galaxies than in the wall dwarf galaxies.  This 
difference is quantified in the K-S test --- the test returned a probability of 
11.1\% that a test statistic greater than or equal to 0.22 will be measured if 
the void sample was drawn from the wall sample.  The void dwarf galaxies have 
lower N/O ratios by an average of $0.05\pm 0.074$ than the wall dwarf galaxies; 
the difference in the median values of the N/O ratio in the void and wall dwarf 
galaxy samples is 0.07.  However, like the shifts seen in the oxygen and 
nitrogen abundances, the shift in the N/O ratio for dwarf galaxies is not 
statistically significant.

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{Images/Paper2/1sig_dwarf_SF_t3_logNO_hist}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/1sig_dwarf_SF_t3_logNO_CDF}
    \caption[N/O distribution of 135 dwarf galaxy sample]{Ratio of nitrogen to 
    oxygen of void dwarf (red solid line) and wall dwarf (black dashed line) 
    galaxies.  A two-sample K-S test of the two data sets results in an 
    asymptotic $p$-value of 0.11, indicating an 11\% probability that a test 
    statistic greater than the observed value of 0.22 will be seen if the void 
    sample was drawn from the wall sample.  This is reflected visually, as the 
    void galaxies appear to have a lower value of N/O than the wall galaxies.  
    This is suggestive of a large-scale environmental influence on the relative 
    chemical abundances in dwarf galaxies.}
    \label{fig:NOratio}
\end{figure*}

%%%% Brighter star-forming galaxies %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We perform the same analysis with the N/O ratio on somewhat brighter galaxies, 
up through $M_r > -20$; the results of this analysis can be seen in Fig. 
\ref{fig:NO_bright} and in Table \ref{tab:stats}.  The shift toward lower N/O 
ratios for the void galaxies is small for all magnitude bins.  The direction of 
the shift between environments for the N/O ratio is consistent for all absolute 
magnitude bins: void galaxies have slightly lower N/O ratios than wall galaxies.  
This is only very weak evidence of a large-scale environmental influence on the 
relative abundances of elements in galaxies, but it is worth testing for in 
larger samples.

% 17-18
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Images/Paper2/1sig_17-20_SF_t3_logNO_stacked}
    \caption[N/O distribution of star-forming galaxies with $-17 > M_r > -20$]
    {Ratio of nitrogen to oxygen of void (red solid line) and wall (black dashed 
    line) star-forming galaxies with $-17 > M_r > -18$ (top), $-18 > M_r > -19$ 
    (middle), and $-19 > M_r > -20$ (bottom).  The results of a two-sample K-S 
    test of the two data sets in each absolute magnitude range can be found in 
    Table \ref{tab:stats}, in addition to other statistics of the samples.  Both 
    the histograms, CDFs, and statistics suggest a very slight difference 
    between the two populations, indicating a mild large-scale environmental 
    influence on the relative chemical abundances of star-forming galaxies with 
    absolute magnitudes $-17 > M_r > -20$.}
    \label{fig:NO_bright}
\end{figure}

%\begin{table}
%
%    \begin{tabu}{cccccccc}
%        Absolute magnitude range & Environment & \# of galaxies & Average & Median & Average shift\footnote[1]{Wall -- Void (Positive shifts indicate that the wall galaxy average is greater than the void average; negative shifts indicate that the void galaxy average is greater than the wall average.)} & $p$-value & KS test statistic \\
%        \hline \\
%        \multirow{2}{*}{Dwarf galaxies} & Void & 42 & $-1.25\pm 0.060$ & -1.28 & \multirow{2}{*}{$0.05\pm 0.074$} & \multirow{2}{*}{0.1116} & \multirow{2}{*}{0.2191}\\
%         & Wall & 89 & $-1.21\pm 0.044$ & -1.22 & & & \\
%        \multirow{2}{*}{$-17>M_r>-18$} & Void & 423 & $-1.08\pm 0.020$ & -1.13 & \multirow{2}{*}{$-0.00\pm 0.024$} & \multirow{2}{*}{0.2657} & \multirow{2}{*}{0.0588}\\
%         & Wall & 895 & $-1.08\pm 0.014$ & -1.09 & & & \\
%        \multirow{2}{*}{$-18>M_r>-19$} & Void & 829 & $-0.86\pm 0.015$ & -0.88 & \multirow{2}{*}{$0.02\pm 0.019$} & \multirow{2}{*}{0.0763} & \multirow{2}{*}{0.0550}\\
%         & Wall & 1498 & $-0.84\pm 0.012$ & -0.85 & & & \\
%        \multirow{2}{*}{$-19>M_r>-20$} & Void & 798 & $-0.62\pm 0.016$ & -0.62 & \multirow{2}{*}{$0.02\pm 0.020$} & \multirow{2}{*}{0.0715} & \multirow{2}{*}{0.0564}\\
%         & Wall & 1490 & $-0.60\pm 0.012$ & -0.60 & & & \\
%    \end{tabu}
%
%    \caption{Statistics of the relative abundance of nitrogen to oxygen in 
%    void and wall galaxies in each of the absolute magnitude ranges listed.  
%    Combined with the histograms in Figs. \ref{fig:NOratio} and 
%    \ref{fig:NO_bright}, these results suggest that there is a large-scale 
%    environmental dependence of the relative abundance of nitrogen to oxygen in 
%    dwarf galaxies.}
%
%    \label{tab:KStest_NO}
%
%\end{table}

Figure \ref{fig:NO_bright} indicates a shift toward higher values in the peak of 
the N/O distribution as the absolute magnitude of the galaxies increases.  There 
is a known positive correlation between the N/O ratio and the stellar mass of a 
galaxy, as discussed in Sec. \ref{sec:Mass_NO} below.  To test whether this 
relation is causing the shift seen in Figs. \ref{fig:NOratio} and 
\ref{fig:NO_bright}, we downsampled the wall galaxies in each magnitude bin to 
match the void sample.  The original shifts in the N/O ratio seen were still 
present after the downsampling; the observed shift in the N/O ratio is not due 
to any variations in the distribution of the stellar masses between the two 
environments.  In addition, if we are overestimating the temperatures in these 
galaxies as a result of an incorrect measurement of [\ion{O}{3}] $\lambda 4363$ 
(discussed above in Section \ref{sec:OH_NH}), that effect should cancel when we 
look at the ratio of nitrogen to oxygen.  This shift toward higher N/O values as 
a function of absolute magnitude indicates that brighter galaxies produce more 
nitrogen than fainter galaxies (relative to their oxygen abundance).  This 
result is consistent with the theory that nitrogen behaves as a secondary 
element in galaxies with high enough metallicity, if we assume a positive 
correlation between absolute magnitude and metallicity.

\begin{sidewaystable}
\centering

	\begin{tabu} to 0.75\textwidth {ccccccccc}
	    Abs. Mag. Range & Environment & \# of Galaxies & Average & Median & Average Shift\footnote[1]{Wall -- Void (positive shifts indicate that the wall values are greater than the void values; negative shifts indicate that the void values are greater than the wall values)} & Median Shift\footnote[1] & K-S Test Statistic & $p$-value \\
	    \hline
	    \hline
        % Oxygen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \multicolumn{9}{c}{\OH}\\
        \hline
        \multirow{2}{*}{Dwarf galaxies} & Void & 42 & $7.99\pm 0.049$ & 8.04 & \multirow{2}{*}{$-0.07\pm 0.060$} & \multirow{2}{*}{-0.03} & \multirow{2}{*}{0.1322} & \multirow{2}{*}{0.6701}\\
         & Wall & 89 & $7.93\pm 0.036$ & 8.01 & & & & \\
        \multirow{2}{*}{$-17>M_r>-18$} & Void & 423 & $7.87\pm 0.016$ & 7.96 & \multirow{2}{*}{$0.01\pm 0.020$} & \multirow{2}{*}{-0.02} & \multirow{2}{*}{0.0455} & \multirow{2}{*}{0.5799}\\
         & Wall & 895 & $7.88\pm 0.012$ & 7.94 & & & & \\
        \multirow{2}{*}{$-18>M_r>-19$} & Void & 829 & $7.74\pm 0.013$ & 7.76 & \multirow{2}{*}{$-0.01\pm 0.016$} & \multirow{2}{*}{-0.04} & \multirow{2}{*}{0.0385} & \multirow{2}{*}{0.4000}\\
         & Wall & 1498 & $7.73\pm 0.009$ & 7.73 & & & & \\
        \multirow{2}{*}{$-19>M_r>-20$} & Void & 798 & $7.59\pm 0.013$ & 7.55 & \multirow{2}{*}{$-0.04\pm 0.017$} & \multirow{2}{*}{-0.05} & \multirow{2}{*}{0.0741} & \multirow{2}{*}{0.0062}\\
         & Wall & 1490 & $7.55\pm 0.010$ & 7.50 & & & & \\
        \hline
        % Nitrogen %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \multicolumn{9}{c}{\NH}\\
        \hline
        \multirow{2}{*}{Dwarf galaxies} & Void & 42 & $6.74\pm 0.035$ & 6.77 & \multirow{2}{*}{$-0.02\pm 0.043$} & \multirow{2}{*}{-0.01} & \multirow{2}{*}{0.1129} & \multirow{2}{*}{0.8383}\\
         & Wall & 89 & $6.72\pm 0.025$ & 6.75 & & & & \\
        \multirow{2}{*}{$-17>M_r>-18$} & Void & 423 & $6.80\pm 0.011$ & 6.82 & \multirow{2}{*}{$0.00\pm 0.014$} & \multirow{2}{*}{-0.01} & \multirow{2}{*}{0.0401} & \multirow{2}{*}{0.7364}\\
         & Wall & 895 & $6.80\pm 0.008$ & 6.81 & & & & \\
        \multirow{2}{*}{$-18>M_r>-19$} & Void & 829 & $6.87\pm 0.009$ & 6.85 & \multirow{2}{*}{$0.01\pm 0.011$} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.0538} & \multirow{2}{*}{0.0873}\\
         & Wall & 1498 & $6.89\pm 0.007$ & 6.87 & & & & \\
        \multirow{2}{*}{$-19>M_r>-20$} & Void & 798 & $6.97\pm 0.009$ & 6.94 & \multirow{2}{*}{$-0.02\pm 0.011$} & \multirow{2}{*}{-0.02} & \multirow{2}{*}{0.0455} & \multirow{2}{*}{0.2272}\\
         & Wall & 1490 & $6.95\pm 0.007$ & 6.93 & & & & \\
        \hline
        % N/O %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \multicolumn{9}{c}{\NO}\\
        \hline
        \multirow{2}{*}{Dwarf galaxies} & Void & 42 & $-1.25\pm 0.060$ & -1.28 & \multirow{2}{*}{$0.05\pm 0.074$} & \multirow{2}{*}{0.07} & \multirow{2}{*}{0.2191} & \multirow{2}{*}{0.1116}\\
         & Wall & 89 & $-1.21\pm 0.044$ & -1.22 & & & & \\
        \multirow{2}{*}{$-17>M_r>-18$} & Void & 423 & $-1.08\pm 0.020$ & -1.13 & \multirow{2}{*}{$-0.00\pm 0.024$} & \multirow{2}{*}{0.04} & \multirow{2}{*}{0.0588} & \multirow{2}{*}{0.2657}\\
         & Wall & 895 & $-1.08\pm 0.014$ & -1.09 & & & & \\
        \multirow{2}{*}{$-18>M_r>-19$} & Void & 829 & $-0.86\pm 0.015$ & -0.88 & \multirow{2}{*}{$0.02\pm 0.019$} & \multirow{2}{*}{0.03} & \multirow{2}{*}{0.0550} & \multirow{2}{*}{0.0763}\\
         & Wall & 1498 & $-0.84\pm 0.012$ & -0.85 & & & & \\
        \multirow{2}{*}{$-19>M_r>-20$} & Void & 798 & $-0.62\pm 0.016$ & -0.62 & \multirow{2}{*}{$0.02\pm 0.020$} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.0564} & \multirow{2}{*}{0.0715}\\
         & Wall & 1490 & $-0.60\pm 0.012$ & -0.60 & & & & \\
         \hline
	\end{tabu}

    \caption[Abundance statistics]{Statistics of the gas-phase oxygen, nitrogen, 
    and nitrogen relative to oxygen abundances in void and wall galaxies in each 
    of the absolute magnitude ranges listed.  Most of these results are not 
    statistically significant, as shown in Figs. 
    \ref{fig:met1sig}--\ref{fig:NO_bright}.  However, the shifts in chemical 
    abundances between the two environments are predominately in the same 
    direction for each of the magnitude bins, suggesting that there is some 
    influence on the chemical evolution of galaxies by the large-scale 
    environment.   Void galaxies have slightly higher oxygen and nitrogen 
    abundances than wall galaxies, but void galaxies have slightly lower N/O 
    ratios than wall galaxies.}

	\label{tab:stats}

\end{sidewaystable}

%-------------------------------------------------------------------------------
\subsection{N/O versus O/H}

Comparing the N/O ratio with the gas-phase oxygen abundance in a galaxy can help 
us understand the nucleosynthesis of nitrogen in galaxies.  When the metallicity 
of a galaxy is low, stars created from this gas do not have enough carbon to 
efficiently produce helium via the CNO cycle.  As a result, any nitrogen 
produced in these stars will behave as a primary element --- it will be produced 
in the same relative quantity as oxygen.  However, when the metallicity of a 
galaxy is high enough, stars are created with enough seed carbon to initiate the 
CNO cycle at an earlier stage in the star's life.  As a result, nitrogen will 
behave as a secondary element and will be produced in a larger quantity relative 
to the primary elements (like oxygen and carbon).  By studying the relation 
between N/O and the metallicity of a galaxy, we should be able to discern the 
critical metallicity at which nitrogen switches from a primary to a secondary 
element.

% Compare with results from Andrews13, Amorin10, VilaCostas93, Lee04, vanZee06, Molla06, Liang06, Pilyugin04, Nava06, Berg12, Thuan95, Henry00, Nicholls14b, PerezMontero09, Pilyugin02, Shields91, Contini02
Our results for N/O versus metallicity can be seen in Fig. \ref{fig:OvN}.  
Unlike many previous comparisons of N/O and metallicity 
\citep[for example,][]{VilaCostas93, Thuan95, Henry00, Pilyugin02, Lee04, 
Pilyugin04, Nava06, vanZee06a, PerezMontero09, Amorin10, Berg12}, we do not find 
a constant value for N/O as a function of O/H for dwarf galaxies (nor for any of 
the somewhat brighter galaxies).  \cite{Shields91}, \cite{Contini02}, and 
\cite{Nicholls14b} also find little or no evidence of a plateau in their study.  
Instead of a constant value for N/O as a function of O/H at low metallicities, 
we find a slight decrease in the N/O ratio as the metallicity increases; a 
linear fit to the dwarf galaxies reveals a slope of $-0.38\pm 0.078$.  This is 
close to the footnoted results of \cite{Andrews13}, who find a slope of $-0.21$ 
for their stellar-mass-binned galaxies with metallicities 
$12 + \log (\text{O}/\text{H}) < 8.5$.  The average value of \NO for the void 
dwarf galaxies is $-1.25\pm 0.060$ with a median value of $-1.28$, while the 
average value for the wall dwarf galaxies is $-1.21\pm 0.044$ with a median of 
$-1.22$.  As shown in Fig. \ref{fig:NOratio}, the void dwarf galaxies have 
slightly less nitrogen relative to oxygen than do dwarf galaxies in denser 
regions.  Both these median values are higher than that of \cite{Andrews13}, and 
these average values are higher than that of \cite{Izotov99} and \cite{Nava06}.

% Perhaps also show the dwarf galaxies colored by SFR for scatter?

\begin{figure}
    \includegraphics[width=0.5\textwidth]{Images/Paper2/1sig_I06_dwarf_0-20_SF_t3_Z12logOH_logNO}
    \caption[N/O versus metallicity of 135 dwarf galaxy sample]{N/O as a 
    function of O/H for star-forming void (red circles) and wall (black 
    triangles) dwarf galaxies.  Error bars have been omitted for clarity.  N/O 
    is expected to be constant for $12 + \log (\text{O}/\text{H}) \lesssim 8.5$, 
    as the metallicity of a galaxy's ISM is too low for stars to be created with 
    enough of the heavy elements to undergo the CNO cycle early enough in their 
    lifetimes.  As a result, nitrogen behaves as a primary element at galactic 
    metallicities less than approximately 8.5.  To place the dwarf galaxies in 
    the context of the general galaxy population, we also plot (gray stars) all 
    star-forming galaxies fainter than $M_r > -20$.}
    \label{fig:OvN}
\end{figure}

If a plateau in the O/H--N/O relation exists, then we should see a slope of 1 in 
the O/H--N/H relation.  When looking at N/H as a function of O/H in Fig. 
\ref{fig:OHvNH}, we see that there is a correlation between the nitrogen and 
oxygen abundances.  However, a best fit to the dwarf galaxies reveals a slope of 
only $0.62\pm 0.078$ --- the nitrogen abundance increases at a slower rate than 
the oxygen abundance.  This result matches the negative relationship between the 
metallicity and the N/O ratio seen in Fig. \ref{fig:OvN}.  If we examine only 
the low-metallicity ($12 + \log(\text{O}/\text{H}) < 7.6$) star-forming galaxies 
with $M_r > -20$, a linear fit produces a slope of $0.05\pm 0.019$ in Fig. 
\ref{fig:OHvNH} and a slope of $-0.94\pm 0.019$ in Fig. \ref{fig:OvN}.  This is 
in sharp contrast to the star-forming galaxies with $M_r > -20$ that have 
metallicities $12 + \log(\text{O}/\text{H}) > 7.6$, where their slope in Fig. 
\ref{fig:OHvNH} is $0.60\pm 0.022$ and $-0.39\pm 0.023$ in Fig. \ref{fig:OvN}.  
It appears that the nitrogen production is independent of the amount of oxygen 
produced in low-metallicity systems.  At normal metallicities 
($7.6 < 12 + \log(\text{O}/\text{H}) < 8.5$), there exists a positive 
relationship between the production of nitrogen and oxygen, although the ratio 
of N/O produced depends on the galaxy's metallicity.

There is no difference between void and wall galaxies in the relationship of 
oxygen and nitrogen production in the low-metallicity sample.  There is a slight 
difference in slopes between the void and wall galaxies with normal 
metallicities, where the void galaxies have a larger slope in the relationship 
between O/H and N/H and a smaller slope in the relationship between O/H and N/O.  
While statistically significant, the difference in the slopes between the two 
environments is not large enough to be physically relevant.  The significant 
scatter in both Figs. \ref{fig:OHvNH} and \ref{fig:OvN} indicates that the 
described relationships between the production of nitrogen and oxygen are only 
global trends in the nucleosynthesis of the galaxies.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper2/1sig_I06_dwarf_0-20_SF_t3_Z12logOH_N12logNH}
    \caption[N/H versus metallicity for 135 dwarf galaxy sample]{N/H as a 
    function of O/H for star-forming void (red circles) and wall (black 
    triangles) dwarf galaxies.  Error bars have been omitted for clarity.  There 
    is a positive correlation between the two abundances.  With a best-fit slope 
    less than 1, we see that the synthesis of nitrogen in these galaxies is 
    primary.}
    \label{fig:OHvNH}
\end{figure}

%-------------------------------------------------------------------------------
%\subsection{Comparison to previously published abundance estimates}

% Amorin10 does not publish their N/O estimates

%-------------------------------------------------------------------------------
\subsection{Mass--N/O relation}\label{sec:Mass_NO}

Just as there is a well-known mass--metallicity relation for galaxies 
\citep[where the metallicity increases with stellar mass; see, 
e.g.,][]{Tremonti04}, there is also a mass--N/O relation.  We expect to see a 
primary N/O plateau in the mass--N/O relation, since galaxies with lower stellar 
masses have not yet produced enough heavy elements to synthesize more nitrogen 
than oxygen.  Beyond the low-mass limit, there should be a steady increase in 
the N/O ratio as a function of stellar mass, due to secondary nitrogen 
enrichment.  Our dwarf galaxies in Fig. \ref{fig:MNO} show a steady increase in 
N/O as a function of stellar mass; there is a hint of the beginnings of a 
plateau for $\log(M_*/M_{\odot}) \lesssim 8$.  The lack of a plateau here could 
be a result of our limited stellar mass range for the dwarf galaxies.  A linear 
fit to our dwarf galaxies reveals a slope of $0.6\pm 0.12$, which is much 
stronger than the slope of $0.30$ found by \cite{Andrews13}.

% Compare with results from Andrews13, Amorin10, PerezMontero09, PerezMontero13
From Fig. \ref{fig:MNO}, we conclude that the N/O plateau, if one exists, starts 
around $\log(M_*/M_{\odot}) \approx 8$.  This is at a much lower mass than that 
found by \cite{Andrews13} --- they claim the N/O plateau exists for galaxies 
with $\log(M_*/M_{\odot}) < 8.9$.  However, our relationship between stellar 
mass and N/O matches Fig. 3 of \cite{Amorin10}, as well as the results of 
\cite{PerezMontero09} and \cite{PerezMontero13}.  None of our data samples 
display an obvious N/O plateau above a stellar mass 
$\log(M_*/M_{\odot}) > 8$, indicative of primary nitrogen versus secondary 
nitrogen production in galaxies.  This could indicate that the switch from 
primary to secondary nitrogen production occurs at a much lower stellar mass 
than found by \cite{Andrews13}.  More low-mass galaxies are needed to extend 
this relation below $\log(M_*/M_{\odot}) < 8$.

\begin{figure}
    \includegraphics[width=0.5\textwidth]{Images/Paper2/MNO_1sig_I06_dwarf_0-20_SF_t3}
    \caption[Stellar mass versus N/O for 135 dwarf galaxy sample]{Mass--N/O 
    relation for star-forming void (red circles) and wall (black triangles) 
    dwarf galaxies.  Error bars have been omitted for clarity.  N/O is expected 
    to remain constant for low stellar masses and increase steadily for larger 
    masses, due to the mass-metallicity relation and the primary vs. secondary 
    synthesis of nitrogen.  To place the dwarf galaxies in the context of the 
    general galaxy population, we also plot (gray stars) the star-forming 
    galaxies with $M_r > -20$.}
    \label{fig:MNO}
\end{figure}

%-------------------------------------------------------------------------------
\subsection{Color--N/O relation}

% Compare to Fig. 9 of vanZee06a, Fig. 6 of Berg12
As \cite{vanZee06a} and \cite{Berg12} discuss, a time delay between the release 
of nitrogen and oxygen will result in a positive relationship between the N/O 
ratio and the color of a galaxy.  If oxygen is primarily produced in higher-mass 
stars (and since these stars die earlier than the intermediate-mass stars 
responsible for the synthesis of nitrogen), then, for a given star formation 
episode, the oxygen produced will be released on a shorter time scale than the 
nitrogen.  As a result, the amount of nitrogen relative to oxygen should 
increase as the hotter, more massive stars die and the galaxy becomes redder.  
This is exactly what we see in Fig. \ref{fig:color_NO}.  We use rest-frame 
colors $K$-corrected to a redshift of 0.1; they are corrected for galactic 
extinction and calculated with model magnitudes \citep{Choi10}.  
\cite{vanZee06a} and \cite{Berg12} also found an increase in the N/O ratio as a 
function of color.

\begin{figure*}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/ur_NO_1sig_I06_dwarf_0-20_SF_t3}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/gr_NO_1sig_I06_dwarf_0-20_SF_t3}
    \caption[Color versus N/O for 135 dwarf galaxy sample]{Color ($u-r$ and 
    $g-r$) vs. N/O ratio for star-forming void (red circles) and wall (black 
    triangles) dwarf galaxies.  Error bars have been omitted for clarity.  N/O 
    is expected to increase as galaxies become redder if there is a time delay 
    between the release of oxygen and nitrogen.  To place the dwarf galaxies in 
    context, we also plot the star-forming galaxies with $M_r > -20$ 
    (gray stars).}
    \label{fig:color_NO}
\end{figure*}

%-------------------------------------------------------------------------------
\subsection{(s)SFR--N/O relation}

We expect there to be a correlation between SFR or sSFR (per unit stellar mass) 
and the N/O ratio in galaxies as a result of the positive correlations between 
the SFR and stellar mass of a galaxy \citep{Brinchmann04} and the sSFR and the 
color of a galaxy (Fig. \ref{fig:color_sSFR}).  As shown in Fig. \ref{fig:MNO}, 
the N/O ratio increases with increasing stellar mass.  As a result, we expect 
there to be a positive correlation between the SFR and N/O ratio in our 
galaxies.  This can be seen in the sample of star-forming galaxies with 
$M_r > -20$ in the left panel of Fig. \ref{fig:SFR_NO}.  Due to the large 
scatter in this relation, the blue star-forming dwarf galaxies exhibit a 
negative correlation between their N/O ratios and their SFRs.  These SFR are 
aperture corrected to estimate the total SFR in the galaxy (not just within the 
SDSS fiber).

\begin{figure*}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/u-r_sSFR_dwarf_0-20_SF_t3}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/g-r_sSFR_dwarf_0-20_SF_t3}
    \caption[Color versus sSFR for 135 dwarf galaxy sample]{Color ($u-r$ and 
    $g-r$) vs. sSFR for star-forming void (red circles) and wall (black 
    triangles) dwarf galaxies.  It is obvious that there is a negative 
    correlation between the $\log (sSFR)$ and the color of a galaxy.  To place 
    the dwarf galaxies in context, we also show the star-forming galaxies with 
    $M_r > -20$ (gray stars).}
    \label{fig:color_sSFR}
\end{figure*}

The N/O ratio is expected to decrease as sSFR increases in galaxies, as is shown 
in the right panel of Fig. \ref{fig:SFR_NO}.  Bluer galaxies have higher sSFR.  
Fig. \ref{fig:color_NO} shows that there is a positive correlation between the 
color and the N/O ratio, such that bluer galaxies have lower N/O ratios.  As a 
result, we are not surprised to see that the N/O ratio decreases as the sSFR 
increases.
%Oxygen is produced in higher-mass stars 
%than nitrogen.  Since a star's lifetime is inversely proportional to its mass, 
%the stars that produce oxygen will die and release their heavy elements to the 
%ISM sooner than the stars responsible for producing nitrogen.  As a result, 
%after a recent star formation episode (when the sSFR is high), the amount of 
%oxygen will increase relative to nitrogen, decreasing the N/O ratio.  As the 
%sSFR of a galaxy declines, its N/O ratio will slowly increase as the nitrogen in 
%the intermediate-mass stars is released.

\begin{figure*}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/SFR_NO_1sig_I06_dwarf_0-20_SF_t3}
    \includegraphics[width=0.49\textwidth]{Images/Paper2/sSFR_NO_1sig_I06_dwarf_0-20_SF_t3}
    \caption[(s)SFR versus N/O for 135 dwarf galaxy sample]{SFR--N/O and 
    sSFR--N/O relations for star-forming void (red circles) and wall (black 
    triangles) dwarf galaxies.  Error bars have been omitted for clarity.  Based 
    on the correlations between SFR and stellar mass and sSFR and color, we 
    expect the N/O ratio to increase with increasing SFR and decrease with sSFR.  
    To place the dwarf galaxies in context, we also show the star-forming 
    galaxies with $M_r > -20$ (gray stars).}
    \label{fig:SFR_NO}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    DISCUSSION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Discussion]{Large-scale environmental influence}\label{sec:environment}

The majority of the shifts in the gas-phase abundances of oxygen, nitrogen, and 
the N/O ratio seen in galaxies fainter than $L_*$ are small and statistically 
insignificant.  However, they occur in almost the same direction across all 
magnitude bins.  This trend suggests that the gas-phase abundances may be 
influenced by the galaxies' large-scale environments.  \cite{Shields91} find no 
offset in the N/O ratio between cluster and field galaxies, despite the 
difference in O/H they observe.  Similar to us, \cite{Contini02} and 
\cite{Pilyugin02} also find a statistically insignificant shift in the N/O ratio 
of cluster galaxies, although they find that these galaxies have lower N/O 
ratios than field spiral galaxies.  Based on Fig. \ref{fig:NOratio} and the 
statistics in Table \ref{tab:stats}, we find weak evidence that void dwarf 
galaxies have a smaller N/O ratio than dwarf galaxies in denser regions.  This 
means that void dwarf galaxies may have more oxygen than wall dwarf galaxies, 
and/or void dwarf galaxies could have less nitrogen than wall dwarf galaxies.  
Here we discuss these possibilities and explore their implications for the 
large-scale environmental impact on the formation and evolution of galaxies.

%If there was more oxygen in void dwarf galaxies than wall dwarf galaxies, then 
%we would expect to see this reflected in the oxygen abundance distributions.  
%However, as Fig. \ref{fig:met1sig} shows, this is not the case --- both void and 
%wall dwarf galaxies have equal distributions of oxygen abundance.  Therefore, to 
%have a difference in oxygen but to return the results seen in Fig. 
%\ref{fig:met1sig}, void dwarf galaxies must then also have more hydrogen than 
%wall dwarf galaxies.  This is definitely possible, as we expect that void 
%galaxies are more privy to an inflow of pristine hydrogen-rich gas than wall 
%galaxies.  However, explaining the excess oxygen in void galaxies is more 
%difficult.  Void dwarf galaxies might retain more of their metals than do wall 
%dwarf galaxies because of the absence of tidal forces, ram pressure stripping, 
%etc. in the void environment.
%
%On the other hand, the difference in Fig. \ref{fig:NOratio} could also be 
%explained by there being less nitrogen in void dwarf galaxies than wall dwarf 
%galaxies.  This would seem to match the theory more, since it is thought that 
%void galaxies should produce fewer metals.  However, this result is not 
%reflected in Fig. \ref{fig:N_1sig}, since the two populations are centered on 
%similar values of $12 + \log \left( \text{N}/\text{H} \right)$.

Table \ref{tab:stats} suggests a slight large-scale environmental dependence of 
the oxygen and nitrogen abundances (relative to hydrogen), where void galaxies 
have slightly more O/H and N/H than wall galaxies.  This small difference is 
not apparent when looking at Figs. \ref{fig:met1sig} and \ref{fig:N_1sig}.  
However, the N/O ratio amplifies this large-scale environmental effect so that a 
shift in the mean (or median) of the two populations can be seen in Fig. 
\ref{fig:NOratio}.  We hesitate to combine the results across all magnitude bins 
in an effort to improve their significance.  Instead, we look toward the future 
to analyze a larger sample of galaxies to increase the statistical significance 
of these results.

\subsection{Higher metallicities in void dwarf galaxies}

A slightly higher metallicity in void dwarf galaxies than wall dwarf galaxies 
may be evidence of a difference in the ratio of dark matter halo mass to stellar 
mass between the two environments.  Simulations by \cite{Jung14} and 
\cite{Tonnesen15} show that the dark matter halo masses of void central galaxies 
are larger than those of wall central galaxies for a given stellar mass.  Due to 
their environment, void dwarf galaxies are more likely to be in the center of 
their own dark matter halo.  Wall dwarf galaxies, on the other hand, are much 
more likely to be a satellite galaxy within a much larger dark matter halo; the 
simulation results mentioned above would not apply to these wall dwarf galaxies.  
However, because the wall dwarf galaxies studied here have sufficiently high 
sSFRs, they most likely live in a small-scale environment very similar to that 
of the void dwarf galaxies, as discussed in Paper I.  As a result, it is likely 
that (and should be tested to see whether) the wall dwarf galaxies in this study 
are actually central galaxies.

Applying the results of these simulations to our dwarf galaxy sample, if the 
ratio of dark matter halo mass to stellar mass is larger in void galaxies, the 
metals ejected from a void galaxy's ISM into its circumgalactic medium are more 
likely to fall back onto the ISM than in a wall galaxy with the same stellar 
mass, since the void galaxy's virial radius and potential well are larger.  As a 
result, two dwarf galaxies with the same stellar mass in these two different 
large-scale environments can have different metallicities --- void dwarf 
galaxies would have higher metallicities than wall dwarf galaxies, matching what 
we see in Table \ref{tab:stats}.

\subsection{Lower N/O ratios in void dwarf galaxies}

A difference in the N/O ratio between void dwarf galaxies and wall dwarf 
galaxies could be a result of the difference in the synthesis of nitrogen in 
galaxies within these two large-scale environments.  If void galaxies are 
retarded in their star formation (as simulations of the $\Lambda$CDM cosmology 
suggest), then cosmic downsizing would reduce the SFR at late times much more in 
wall galaxies than in void galaxies.  As a result, the minimum gas-phase 
metallicity required for the production of secondary nitrogen in walls would be 
achieved at an earlier time in the galaxy's lifetime than in a void galaxy.  
This would cause the N/O ratio in wall galaxies to be larger than that in voids.  
\cite{vanZee06a} suggest that a galaxy with a declining SFR (wall galaxies) will 
have a higher nitrogen-to-oxygen yield than a galaxy with a constant SFR (void 
galaxies).  This is due to more oxygen being released into the ISM as a result 
of the ongoing star formation in the void galaxies.  This explanation is 
supported by the color--N/O diagram in Fig. \ref{fig:color_NO}, which reveals 
that redder galaxies have higher N/O ratios.  The correlation between color and 
the N/O ratio found in \cite{vanZee06a}, \cite{Berg12}, and Fig. 
\ref{fig:color_NO} is a result of declining SFRs \citep{vanZee06a}.  Therefore, 
the shift in the N/O ratio we see between void and wall galaxies may be 
observational evidence of retarded star formation in void galaxies as a result 
of cosmic downsizing.
%If wall galaxies have a higher 
%N/H ratio than void galaxies, they could be producing nitrogen as a secondary 
%element more so than void dwarf galaxies.  This would cause the N/O ratio in 
%wall galaxies to be larger than that in voids.

Another explanation that would lead to a shift in the N/O ratio between 
environments is a difference in the ratio of intermediate- and high-mass stars 
produced in void and wall dwarf galaxies.  For there to be more oxygen relative 
to nitrogen in void dwarf galaxies, the percent of higher-mass stars produced in 
a star formation episode would be higher than that in wall dwarf galaxies.  This 
would indicate a varying IMF as a function of large-scale environment.  Previous 
studies have been inconclusive when testing for a varying IMF \citep[see][for 
example]{Kroupa01, Kroupa02, Hoversten08, Meurer09}.  It is beyond the scope of 
this paper to elaborate on this explanation.
%This is the opposite of the (extremely mild) offset between the two environments 
%that Ginny found.

%-------------------------------------------------------------------------------
\subsection{N/O ratio for extremely low metallicity galaxies}

While Paper I shows that there is not a special population of extremely low 
metallicity dwarf galaxies residing in voids, we want to look in particular at 
the N/O ratio of extremely low metallicity galaxies.  For the 21 dwarf galaxies 
with $12 + \log (\text{O}/\text{H}) < 7.6$ identified in Paper I, we see from 
Fig. \ref{fig:OHvNH} that their N/H ratios are also some of the lowest in the 
dwarf galaxy sample.  However, as shown in Fig. \ref{fig:OvN}, their N/O ratios 
cover the range of N/O ratio values of all the dwarf galaxies studied.  This is 
consistent with the expectation that nitrogen behaves as a primary element for 
galaxies with low and moderate metallicities.  Details of these 21 dwarf 
galaxies with extremely low metallicities are listed in Table \ref{tab:lowZ_P2}, 
including their gas-phase chemical abundances.

\input{Tables/Paper2/XMPtable}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%    CONCLUSION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The nucleosynthesis of nitrogen is a vital component of the chemical evolution 
of galaxies in our universe.  We estimate the nitrogen abundance and N/O ratio 
of dwarf galaxies using the direct $T_e$ method and spectroscopic line flux 
measurements from the SDSS DR7 sample as reprocessed in the MPA-JHU catalog.  
The 135 galaxies analyzed suggest a slight large-scale environmental dependence 
of the N/O ratio, where void dwarf galaxies could have a lower N/O ratio than 
dwarf galaxies in denser environments.  Thus, the large-scale 
($\sim 10\text{ Mpc}$) environment might influence the chemical evolution of 
dwarf galaxies.

We find small, statistically insignificant shifts in the mean (or median) N/O 
ratio for galaxies between the void and denser regions across all blue, 
star-forming galaxies with $M_r > -20$.  These shifts are somewhat more 
significant, however, when we look at the entire sample of galaxies.  Each 
magnitude bin is shifted in the same direction, and are potentially very 
interesting, as they might indicate delayed star formation histories, more 
constant SFRs, and larger ratios of dark-matter-halo-mass to stellar mass in 
void galaxies, as discussed in Section \ref{sec:environment}.  A larger sample 
would help test these results.  We look to increase the sample and probe a 
larger magnitude (and mass) range of dwarf galaxies in Douglass \& Vogeley 
(2017, in preparation).

In addition, we look at the relationship between the N/O ratio and other 
physical characteristics of our dwarf galaxies.  In the relation between N/O and 
O/H, our galaxies all reside on the so-called ``nitrogen plateau,'' where the 
N/O ratio is predicted to be constant for low and intermediate metallicities.  
However, instead of a constant value for these galaxies, we find a negative 
correlation between the N/O and O/H ratios.  Our dwarf galaxies show a positive 
correlation between stellar mass and the N/O ratio.  These dwarf galaxies have 
some of the lowest N/O ratios for both their color and (s)SFR.  Beyond the 
suggestive large-scale environmental dependence of the N/O ratio, there is no 
clear large-scale environmental dependence in any of these relationships.

The N/O ratios of the extremely metal-poor dwarf galaxies are no different than 
those of the remaining dwarf galaxy sample, though their N/H abundance is also 
extremely low.  A more detailed study of these 21 extremely metal-poor dwarf 
galaxies is recommended to confirm their abundance values and discover any 
characteristics shared by the population.

Although SDSS provides spectroscopic observations for over 800,000 galaxies, 
only 135 are dwarf galaxies with emission-line fluxes necessary to estimate the 
gas-phase chemical abundances using the direct $T_e$ method.  The greatest 
limiting factor in this sample is the requirement of the [\ion{O}{2}] 
$\lambda 3727$ doublet in the abundance calculations.  We seek to develop a 
work-around for this emission line in Douglass \& Vogeley (2017, in preparation) 
to greatly increase our sample of dwarf galaxies with abundance estimates.  
These estimated ionic abundances can then be compared with environmental 
dependence of star formation and abundance predictions from high-resolution 
hydrodynamic simulations.

Further tests may refine our understanding of the environmental scale that is 
important for determining the chemical evolution of dwarf galaxies.  In 
particular, it will be important to examine whether the influence of relatively 
small-scale ($\sim$2 Mpc) environments is more significant to a dwarf galaxy's 
chemical evolution than the larger-scale environment investigated here.  In 
previous work, both \cite{Kreckel15} and \cite{Beygu17} find little evidence to 
support a significant large-scale environmental influence on the gas content, 
chemical content, or SFR of void galaxies.  Future work will expand on these 
studies with a larger sample and the possible influence they might have on the 
dwarf galaxies' chemical contents and SFRs.  \cite{Beygu17} also investigate any 
connection between a galaxy's physical properties and its location within a 
void.  We also look to study this possible connection with the increased sample 
size of dwarf galaxies in Douglass \& Vogeley (2017, in preparation).