### 2017-06-02 KellyDouglass/Thesis/Thesis_sections/Paper3.tex

\chapter{Large-scale environmental dependence of chemical abundances in dwarf galaxies and implications for connecting star formation and halo mass}