### 2017-06-05 KellyDouglass/Thesis/Thesis_sections/Conclusion.tex
Small-scale theory, data

\chapter{Conclusions}

\section{Suggestions for future work}