### 2017-06-09 KellyDouglass/Thesis/preamble.tex

\begin{preamble}

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dedications
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{DUTdedications}
\begin{center}
Dedications
\end{center}
\end{DUTdedications}

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgments}

% Advisor

% Thesis committee and collaborators (Renyue, Daniele, Jinfu, Brean, Salvatore)
I would like to thank our collaborator Renyue Cen for his help in the 
interpretation of these results.  I would also like to acknowledge Crystal 
Moorman for her help and support throughout this work.  Finally, I would like to 
thank the referees (including Burcu Beygu) for their detailed comments and 
critique of my work.

% Grad students & PLBD

% Family

% Steve

%Support for this work was provided by NSF grant AST--1410525.
%
%Funding for the SDSS and SDSS-II has been provided by the Alfred P. Sloan 
%Foundation, the Participating Institutions, the National Science Foundation, the 
%U.S. Department of Energy, the National Aeronautics and Space Administration, 
%the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education 
%Funding Council for England.  The SDSS Web Site is \emph{http://www.sdss.org/}.
%
%The SDSS is managed by the Astrophysical Research Consortium for the 
%Participating Institutions.  The Participating Institutions are the American 
%Museum of Natural History, Astrophysical Institute Potsdam, University of Basil, 
%University of Cambridge, Case Western Reserve University, University of Chicago, 
%Drexel University, Fermilab, the Institute for Advanced Study, the Japan 
%Participation Group, Johns Hopkins University, the Joint Institute for Nuclear 
%Astrophysics, the Kavli Institute for Particle Astrophysics and Cosmology, the 
%Korean Scientist Group, the Chinese Academy of Sciences (LAMOST), Los Alamos 
%National Laboratory, the Max-Planck-Institute for Astronomy (MPIA), the 
%Max-Planck-Institute for Astrophysics (MPA), New Mexico State University, Ohio 
%State University, University of Pittsburgh, University of Portsmouth, Princeton 
%University, the United States Naval Observatory, and the University of 
%Washington.

\end{acknowledgments}

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of Contents
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List of Tables
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\listoftables

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List of Figures
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\listoffigures

\iffinal{}{\newpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

% Papers 1-3

% Small-scale enivronment

% GV

\end{abstract}

\iffinal{}{\newpage}

\end{preamble}