### 2017-06-10 KellyDouglass/Thesis/Thesis_sections/Paper3.tex

\chapter{Large-scale environmental dependence of chemical abundances in dwarf galaxies and implications for connecting star formation and halo mass}\label{ch:Paper3}

This chapter has been submitted to the \emph{Astrophysical Journal} by Kelly A. 
Douglass, Michael S. Vogeley, and Renyue Cen; it will be referenced as Douglass 
et. al (2017).