### 2017-06-12 KellyDouglass/Thesis/CV.tex
Updated Intro, Conclusion

\newcommand{\sectionrule}{\noindent \hfil\rule{\textwidth}{.6pt}\hfil}

\iffinal{}{\newpage}

\begin{vita}

{\Large\bf Kelly A. Douglass}

\singlespacing

\noindent
{\large\bf {\sc Education}}

\sectionrule
\vspace{1mm}

\begin{description}[noitemsep]
    \item[Drexel University] Philadelphia, Pennsylvania USA
    \begin{description}[noitemsep]
	    \item[] Ph.D., Physics, July 2017
	    \item[] M.S., Physics, June 2013
    \end{description}
    \item[Cornell University] Ithaca, New York USA
    \begin{description}
	    \item[] B.A., Physics, May 2011
    \end{description}
\end{description}

\noindent
{\large\bf {\sc Publications}}

\sectionrule

\begin{itemize} %\itemsep -2pt
    \item Douglass, K.A., Vogeley, M.S., \& Cen, R.  ``Large-scale environmental dependence of chemical abundances in dwarf galaxies and implications for connecting star formation and halo mass". \href{https://doi.org/}{submitted to ApJ}, 2017
	\item Douglass, K.A. \& Vogeley, M.S.  ``Large-scale environmental dependence of the abundance ratio of nitrogen to oxygen in blue, star-forming galaxies fainter than $L_*$". \href{https://doi.org/10.3847/1538-4357/aa5e53}{ApJ, 837: 42--55}, 2017
	\item Douglass, K.A. \& Vogeley, M.S.  ``Determining the large-scale environmental dependence of gas-phase metallicity in dwarf galaxies". \href{https://doi.org/10.3847/1538-4357/834/2/186}{ApJ, 834: 186--198}, 2017
\end{itemize}

\noindent
{\large\bf {\sc Awards}}

\sectionrule

\begin{itemize}[noitemsep]
    \item Drexel University College of Arts \& Sciences Outstanding Dissertation Award, 2017
    \item AAS Chambliss Astronomy Achievement Student Award, 2014
    \item Drexel University Continuing Teaching Excellence Award, 2014
    \item Drexel University Teaching Excellence Award, 2013
    \item Drexel University Teaching Excellence Award --- Highly Commended, 2012
    \item Drexel University Department of Physics First Year Graduate Student Award, 2012
    \item Drexel University College of Arts \& Sciences Dean's Scholarship, 2011--13
\end{itemize}

\noindent
{\large\bf {\sc Teaching Experience}}

\sectionrule

\begin{description}[noitemsep]
    \item[Teaching Assistant] Physics 135: How Things Work, Drexel University
    \item[Teaching Assistant] Physics 113, 114, 115: Contemporary Physics I, II, III, Drexel University
    \item[Teaching Assistant] Physics 102, 201: Fundamentals of Physics II, III, Drexel University
    \item[Teaching Assistant] Education T580: TA Orientation and Preparation, Drexel University
    \item[Teaching Assistant \& Consultant] Computer Science 1112: Introduction to MATLAB, Cornell University
    \item[Grader] Honors 301: Special Theory of Relativity, Drexel University
    \item[Grader] Physics 428: Quantum Mechanics III, Drexel University
\end{description}

\noindent
{\large\bf {\sc Public Outreach}}

\sectionrule

\begin{itemize}[noitemsep]
	\item \textbf{Professor Bob's Hands-On Science} \hfill \textit{June 2015 to July 2017}
	\item \textbf{Volunteer at the Lynch Observatory} \hfill \textit{September 2011 to June 2017}
\end{itemize}

\end{vita}