### 2017-06-12 KellyDouglass/Thesis/Thesis_sections/Introduction.tex

\chapter{Introduction}

% Description of large-scale structure in our universe
%  - What has been observed?
%  - What has been simulated?

\begin{figure}
    \includegraphics[width=0.65\textwidth]{Images/Intro/sdss}
    \caption[SDSS galaxy map]{A slice through the SDSS galaxy distribution; each 
    point on this plot is a galaxy.  Large sky surveys such as SDSS have 
    revealed a non-uniform distribution of galaxies throughout the universe.  
    Dubbed the cosmic web, galaxies clump together in clusters and along 
    filaments, leaving giant gaps between them (similar to a sponge).}
    \label{fig:SDSS_map}
\end{figure}

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Millenium_simulation}
    \caption[Dark matter simulation]{Snapshots of a portion of the Millenium 
    Simulation showing the dark matter distribution at different points in time 
    from 0.21 Gyr after the Big Bang to the present.  Small perturbations in the 
    initial distribution of dark matter are amplified as the universe expands.  
    Gravity causes the slightly overdense regions to contract while 
    simultaneously causing the underdense regions to expand.}
    \label{fig:DMsim}
\end{figure}

Large sky surveys \citep[such as the Sloan Digital Sky Survey --- 
SDSS;][]{York00} have revealed a non-uniform distribution of galaxies throughout 
the universe.  Taking on a shape similar to that of a sponge or a 
three-dimensional cosmic web \citep{Bond96}, galaxies clump together in clusters 
and along filaments, leaving void regions between these large-scale structures.  
Evidence of this distribution can be seen in Fig. \ref{fig:SDSS_map} for 
galaxies in the SDSS Data Release 7 \citep[SDSS DR7;][]{Abazajian09}.  Based on 
these observations, dark matter simulations have been constructed which 
successfully reproduce the same large-scale structure.  If we start with a 
mostly uniform distribution of dark matter at the Big Bang, Fig. \ref{fig:DMsim} 
outlines the evolution of that dark matter up through the present.  These 
snapshots from the Millenium Simulation project \citep{Springel05} show that 
small perturbations in the initial distribution of dark matter are amplified as 
the universe expands.  Gravity causes the slightly overdense regions to collapse 
while simultaneously causing the underdense regions to expand.  Due to dark 
matter's strong gravitational interaction, we believe that the baryonic mattter 
(matter which interacts electromagnetically) will trace the dark matter 
distribution, resulting in the galaxy distribution we observe today.

% Void galaxies
\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/VoidFinder}
    \caption[Sky map highlighting voids and void galaxies]{A 10 \hMpc slice of 
    SDSS DR7 \citep[Fig. 1]{Moorman14} with void regions highlighted in blue 
    circles.  Void galaxies are shown as red points while wall galaxies are 
    black.  Existing in the cosmological voids (space between the galactic 
    filaments), void galaxies are though to demonstrate the fundamental 
    characteristics of galactic evolution.}
\end{figure}

Existing in the cosmological voids, void galaxies are thought to demonstrate the 
fundamental characteristics of galactic evolution.  Cosmic voids are an 
important environment for studying galaxy formation \citep[see][for a review]
{vandeWeygaert11}.  Gravitational clustering proceeds as if in a very low 
density universe, where the amassment of gravitationally bound dark matter halos 
ends relatively early and there is little subsequent interaction between the 
galaxies due to the lower density and the faster local Hubble expansion.  
Therefore, $\Lambda$CDM cosmology predicts that galaxies formed in voids should 
have lower mass and be retarded in their star formation when compared to those 
in denser regions \cite[e.g.,][]{Gottlober03,Goldberg05,Cen11}.  
\cite{Goldberg04} show that a void region that is only 10\% of the mean density 
of the universe with $\Omega_{matter} = 0.3$, $h = 0.7$ dynamically evolves as 
if $\Omega_{matter} = 0.02$, $\Omega_\Lambda = 0.48$, and $h = 0.84$.  
Additionally, hydrodynamical cosmological simulations by \cite{Cen11} show that 
void galaxies continue to form stars because the gas in voids remains below the 
critical entropy threshold.  Void galaxies evolve in a relatively pristine 
environment where interactions are rare and star formation proceeds up to the 
present epoch because void galaxies are able to retain their gas.  This 
contrasts the denser environments, where the chemical composition and evolution 
of galaxies are drastically altered due to mergers and tidal and/or ram-pressure 
stripping.

% What do we already know about the environmental influence?
Observational studies of void galaxies have included investigation of 
photometric properties such as luminosity \citep{Hoyle05,Croton05,Moorman15}, 
color and morphological type \citep{Grogin00,Rojas04,Patiri06,Park07,
vonBendaBeckmann08,Hoyle12}, star formation rates estimated from optical 
spectroscopy and UV photometry \citep{Rojas05,Moorman15,Beygu16}, and gas 
content \citep{Kreckel12,Moorman16,Jones16}.  When compared to galaxies in 
denser regions, void galaxies are fainter, of late morphological types, bluer, 
forming stars at higher rates per unit stellar mass, and more gas rich.

% Dwarf galaxies
\begin{figure}
    \includegraphics[width=0.65\textwidth]{Images/Intro/1sig_13-23_SDSS_Mr_hist_count_fill}
    \caption[Absolute magnitude distribution of galaxies in SDSS DR7]{Absolute 
    magnitude ($M_r$) distribution of galaxies in SDSS DR7, separated into void 
    and wall environments.  Dwarf galaxies are defined as galaxies with absolute 
    magnitudes fainter than -17 ($M_r > -17$).  For reference, the Milky Way has 
    an absolute magnitude $M_r \approx -20$.}
    \label{fig:Mr_SDSS}
\end{figure}

In particular, we study the effects of the large-scale environment on the 
formation and evolution of dwarf galaxies.  Fig. \ref{fig:Mr_SDSS} shows the 
distribution of galaxies by their absolute magnitudes detected in SDSS DR7.  
Defined to be galaxies with absolute magnitudes fainter than -17 ($M_r > -17$), 
the low stellar masses of these systems correspond to more shallow gravitational 
potential wells.  This causes any environmental effects to have a more 
significant influence on dwarf galaxies than on galaxies with larger stellar 
masses.  In addition, studying void dwarf galaxies will complement previous 
studies of dwarfs in groups and clusters because the assembly histories of 
low-mass galaxies are predicted to be very different \citep[e.g.,][]{Gao07,
Lackner12}.  Observations to date also show that the properties of dwarf 
galaxies vary dramatically with the environment \citep[e.g.,][]{Ann08,Geha12}.

% Metallicity
%  - How are elements synthesized in stars?
%  - What does the light look like when it reaches us?  When we look at an emission spectrum, what are we studying?
%  - What are forbidden transitions and why are they important?
%  - How do we get the temperature from the emission line ratio?  (Summary - refer to appendix for more details)
%  - What is the metallicity?  O/H - Why oxygen?  and once we estimate the abundances of each ion, we add them together to get the total abundance for that element

The first stars that formed are thought to have been composed of only hydrogen 
(and possibly some trace amount of helium).  As stars burn, nucleosythesis 
creates heavier elements at the core.  When the stars die and expel their 
content, the heavier elements are distributed between the surrounding 
interstellar gas and circumgalactic medium.  When new stars form from this gas, 
they include both hydrogen and some of the heavier elements.  By measuring the 
fraction of elements heavier than helium (``metals'') relative to the hydrogen 
content (the metallicity), we measure the integrated star formation history of 
the galaxy.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Kirchoff}
    \caption[Kirchoff's laws of spectroscopy]{Three different viewpoints of the 
    light emitted from a star and/or a cool cloud of gas, representing 
    Kirchoff's laws of spectroscopy.  A star emits light across all wavelengths, 
    so the resulting spectrum is known as a continuous spectrum.  Star's light 
    that has passed through a cloud of cool gas before being observed is an 
    absorption spectrum, where the elements in the cloud have absorbed some of 
    the light at specific wavelengths (corresponding to the particular elements 
    present in the gas).  If observed off the line-of-sight of the star, the 
    light emitted from the cloud of cooler gas results in an emission spectrum, 
    where the gas is re-radiating the light it absorbed from the star.}
    \label{fig:Kirchoff}
\end{figure}

We use a galaxy's spectrum to measure various properties of the stellar 
population and interstellar medium.  A galaxy's spectrum is a superposition of 
light from the stars and emission lines from the gas.  Depending on our 
viewpoint of the galaxy's light, we will see a different type of spectrum.  As 
shown in Fig. \ref{fig:Kirchoff}, if we separate a star's light by its 
wavelength, we will observe a continuous spectrum, since it is a blackbody.  If 
instead we observe the starlight after it has passed through a cloud of cooler 
gas, we will measure an absorption line spectrum; the elements in the gas cloud 
have absorbed some of the star's light at specific wavelengths corresponding to 
the particular ions present.  Finally, if we only observe the light emitted from 
the cool gas cloud, we will measure an emission line spectrum.  The method we 
employ to estimate the metallicity of a galaxy requires particular emission 
lines.  We study the gas-phase chemical abundances of \ion{H}{2} regions in a 
galaxy, where the light is coming from low-density gas partially ionized by UV 
photons emitted from young, hot stars.  Therefore, we are measuring the 
abundances of the elements produced in previous generations of stars.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/spectrum}
    \caption[Sample (void) dwarf galaxy spectrum]{Example SDSS DR7 spectrum --- 
    this is a void dwarf galaxy whose gas-phase chemical abundances are slightly 
    above average.  The [\ion{O}{3}] $\lambda$4363 auroral line is circled in 
    red, while the [\ion{O}{3}] $\lambda \lambda$4959,5007 doublet is circled in 
    yellow.}
    \label{fig:spectrum}
\end{figure}

The ion transitions observed in the optical part of the spectrum (where SDSS DR7 
operates) consist of forbidden transitions.  As explained in detail in Section 
\ref{sec:physics}, forbidden transitions have much longer lifetimes than the 
allowed electron transitions.  These transitions depend strongly on the 
temperature and density of the electron gas, because the long lifetimes allow 
the atoms to easily collisionally de-excite.  By taking the ratio of the 
[\ion{O}{3}] $\lambda$4363 auroral line to the [\ion{O}{3}] 
$\lambda \lambda$4959,5007 doublet, we can get an estimate of the temperature of 
the gas.  These lines are highlighted on the example SDSS DR7 spectrum shown in 
Fig. \ref{fig:spectrum}.

Oxygen is used as a proxy to measure the metallicity of an \ion{H}{2} region, 
because oxygen is the most abundant element in the universe after hydrogen and 
helium.  It also has strong emission lines in the visible wavelength range, so 
it is relatively easy to observe.  Oxygen emission is the most effective cooling 
source in an \ion{H}{2} region, so lower metallicities correspond to higher 
temperatures.  There are three main factors which determine the strength of an 
emission line: the temperature, density, and number of ions in the gas.  It is 
possible to calculate the number ratios for any of the heavy elements with 
respect to hydrogen once the electron temperature and density of the 
interstellar gas are estimated.

% Outline of Thesis
% Pose the problem - how does the large-scale environment affect galaxy evolution?
We want to understand how dark matter affects the evolution of a galaxy and what 
influence it has on a galaxy's star formation.  To do this, we compare the 
properties of galaxies living in voids with galaxies in denser regions in an 
effort to understand how the properties of the gas and the history of star 
formation in a galaxy depend on the environment.  We then try to infer what 
these results tell us about the dark matter structure and history of the galaxy.  
In Chapter \ref{ch:Paper1}, we calculate and compare the metallicity of 135 
star-forming dwarf galaxies as a function of their large-scale environment.  We 
continue studying this galaxy sample in Chapter \ref{ch:Paper2} by estimating 
their ratios of nitrogen to oxygen and looking at how that depends on the 
large-scale environment.  To expand our dwarf galaxy sample, we derive an 
approximation for the amount of singly-ionized oxygen and repeat our analysis of 
the large-scale environmental influence on the gas-phase chemical abundances of 
1920 star-forming dwarf galaxies in Chapter \ref{ch:Paper3}.  In Chapter 
\ref{ch:smallScaleEnvironment}, we investigate how the dwarf galaxies' 
properties depend on the galaxies' distances to the nearest neighbors and 
groups.  Finally, in an effort to better understand how galaxies evolve from 
blue, star-forming systems to ``red and dead,'' we discover a way to classify 
galaxies into their location on the color-magnitude diagram in Chapter 
\ref{ch:GV}.