### 2017-06-15 KellyDouglass/Thesis/Thesis_sections/smallScaleEnvironment.tex

\chapter[Small-scale environment]{The influence of the small-scale environment on dwarf galaxy evolution}\label{ch:smallScaleEnvironment}

This work was done in collaboration with Daniele Schneider for her senior thesis 
``The effects of small-scale environment on dwarf galaxies.''

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Introduction]{Introduction}

%% Large-scale environment influence
%Large galaxy redshift surveys have shown that the large-scale structure of the 
%distribution of galaxies in the universe is similar to that of a 
%three-dimensional cosmic web \citep{Bond96}, where galaxy clusters are connected 
%by thin filaments of galaxies and are separated by voids (large, underdense 
%regions that occupy close to 60\% of space).  Over the last few decades, the 
%Sloan Digital Sky Survey \citep{Abazajian09,Ahn12} has accelerated the research 
%into the influence of the large-scale environment on the formation and evolution 
%of galaxies.
%
%Because gravitational clustering within a void proceeds as if in a very 
%low-density universe, cosmic voids are an important environment for studying 
%galaxy formation \citep[see][for a review]{vandeWeygaert11}.  The $\Lambda$CDM 
%cosmology predicts that galaxies formed in voids have lower masses and are 
%retarded in their star formation when compared to those in more dense 
%environments \citep{Gottlober03,Goldberg05,Cen11}.  The effects of the void 
%environment should be most obvious in dwarf galaxies due to their minimal 
%gravitational potential.  As a result, they are more sensitive to astrophysical 
%effects such as cosmological reionization, internal feedback from supernovae and 
%photoheating from star formation, small-scale details of dark matter halo 
%assembly, and the properties of dark matter.
%
%Observations have shown that the properties of dwarf galaxies vary dramatically 
%with the environment \citep[e.g.,][]{Ann08,Geha12}.  Void galaxies have been 
%found to have lower stellar mass \citep{Hoyle05,Croton05,Moorman15}, be bluer 
%and of a later type \citep{Grogin00,Rojas04,Patiri06,Park07,vonBendaBeckmann08,
%Hoyle12}, have higher star formation rates \citep{Rojas05,Moorman15,Beygu16}, 
%and be more gas rich \citep{Kreckel12,Moorman16,Jones16} than galaxies in denser 
%regions.

% Small-scale environment - what has been done so far?
In conjunction with the large-scale environment, the small-scale environment 
($\sim$1 \hMpc) has also been found to influence a galaxy's evolution.  A 
well-established morphology-density relation exists \citep{Dressler80}, where 
the fraction of late-type galaxies is inversely proportional to the local 
density.  \cite{Ellison09} determined that a galaxy's local environment 
influences a galaxy's evolution more than its large-scale environment.  
Likewise, \cite{Rupke08} concluded that interacting galaxies have suppressed 
metallicities, because interactions induce flows of hydrogen.  \cite{Park09} 
found that, for galaxies with $M_r < -19$, galaxy interactions out to the virial 
radius of the nearest neighbor influence the evolution of the target galaxy.  
Along with \cite{Park07}, they determine that the large-scale environment has a 
minimal effect on the evolution of a galaxy once the luminosity and morphology 
are taken into account.

We want to understand if the small-scale environment is more influential in a 
galaxy's evolution than its large-scale environment.  One important question 
within the small-scale environment asks if tidal influences due to the nearest 
neighbor galaxy or gravitational potentials from the nearest galaxy group are 
more influential in a galaxy's evolution.  A group consists of multiple galaxies 
that share the same dark matter halo.  A group member will interact 
gravitationally with all other group members, not just its nearest neighbor.  In 
groups and clusters dense enough to strip and heat the gas, galaxies will also 
experience ram pressure stripping.  And a galaxy near a group, but not within 
it, may also experience strong tidal effects from the group similar to and 
stronger than those from a neighboring galaxy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Nearest neighbor calculations]{Identifying the nearest neighbor}\label{sec:Theory_dist}

We employ two different spatial metrics to determine the best way to locate a 
galaxy's nearest neighbor.  We set a maximum velocity separation of 300 km/s 
between a galaxy and its nearest neighbor to capture the majority of interacting 
galaxies \citep[according to the distribution in peculiar velocity shown in][]
{Hwang10}.  Within a relative velocity $v_{rel} < 300\text{ km/s}$, we use 
either the physically closest galaxy in units of \hMpc or the galaxy within the 
smallest fraction of its virial radius; both these methods are described below 
in further detail.  The two methods result in different neighboring galaxies for 
about 40\% of our dwarf galaxy sample.

\subsection{Peculiar velocity}
% Finger of God

A galaxy's redshift is composed of both the expansion of the universe and the 
galaxy's peculiar velocity (its motion relative to its surrounding galaxies and 
environment).  Known as the ``Finger of God'' effect, the peculiar velocity 
causes galaxy groups and clusters to appear extended along the line of sight 
when using redshift as a distance measurement.  When we calculate distances on 
large scales, the peculiar velocity is negligible.  However, when we study the 
distance between two objects in the universe on a small scale (within a few 
\hMpc), the peculiar velocity dominates.  The relative velocity $v_{rel}$ 
between galaxies $a$ and $b$ is defined as
\begin{equation}
    v_{rel} = |z_a - z_b|c
\end{equation}
where $z$ is the redshift.  We require all 
neighbors to have a maximum relative velocity $v_{rel} < 300\text{ km/s}$, as we 
believe those outside this range will have minimal influence on the target 
galaxy's evolution.  We show in Sec. \ref{sec:selection_effects} that our 
results are insensitive to this value.

\subsection{Sky separation in \hMpc}

The projected distance between two galaxies is found by projecting the two 
galaxies onto a sphere of radius $\bar{r}$, where $\bar{r}$ is the average 
distance from Earth of the two galaxies.  This assumes that two galaxies which 
are gravitationally bound are in a system where the relative velocity between us 
and it is 0 and that $\bar{r}$ is the distance to the center of the system.  For 
redshifts $z \ll 1$, we can approximate the distance to the system is as
\begin{equation}
    \bar{r} = \frac{\bar{z}c}{H_0}
\end{equation}
where $\bar{z}$ is the average redshift of the two galaxies, $c$ is the speed of 
light, and $H_0$ is the Hubble constant.  Each galaxy's right ascension (RA) and 
declination (dec.) are converted to Cartesian coordinates to find their sky 
separation.

\subsection{Fractional virial radii}
% Virial radius calculation

The virial radius of a galaxy defines the region within which something is 
gravitationally bound to the galaxy.  From \cite{Hwang10}, we calculate the 
virial radius of a galaxy to be
\begin{equation}\label{eqn:r_vir}
    r_{vir} = (3L\gamma/4\pi/200\rho_c)^{1/3}
\end{equation}
where $L$ is the galaxy's luminosity, $\gamma$ is the mass-to-light ratio, and 
$\rho_c$ is the critical density of the universe.  The critical density is the 
density at which evenly distributed gas would collapse to form a galaxy.

\cite{Choi07} find that the velocity dispersion of a galaxy depends on its 
morphological type, where $\sqrt{2}\sigma_{late} \approx \sigma_{early}$.  In 
terms of the velocity dispersion, the kinetic energy of a galaxy 
$K = 3/2 \sigma^2$.  Its gravitational potential energy $U = 3GM/5R$.  
Therefore, the virial theorem shows us that
\begin{align}
    3\sigma^2 \simeq \frac{3}{5} \frac{GM}{R}\nonumber \\
    M \simeq \frac{5\sigma^2 R}{G}
\end{align}
With $\gamma \equiv M/L$, at a fixed luminosity and virial radius we see that
\begin{equation}
    \gamma_{late} \simeq \frac{5\sigma_{late}^2 R}{GL} \qquad \gamma_{early} \simeq \frac{5\sigma_{early}^2 R}{GL}
\end{equation}
Taking the ratio of the two mass-to-light ratios and using the relation between 
the velocity dispersions found by \cite{Choi07}, we find that the mass-to-light 
ratio depends on the galaxy's morphology type: $\gamma_{early} = 2\gamma_{late}$.

\cite{Hwang10} defines $200\rho_c = 740\bar{\rho}$, where 
$\bar{\rho} = (0.0223\pm 0.0005)(\gamma_{late}L_{-20})(\text{\hMpc})^3$.  We can 
then rewrite Eqn. \ref{eqn:r_vir} as
\begin{equation}
    r_{vir} = \left( \frac{3}{4\times 740\times 0.0223\pi} \frac{\gamma}{\gamma_{late}} \frac{L}{L_{-20}} \right)^{1/3}
\end{equation}
where $L_{-20}$ is the luminosity of a galaxy with $M_r = -20$.  We can then use 
the virial radius to scale the distance between two galaxies as a fraction of 
the neighbor's virial radius.

Rather than using a measure of the virial radius for the group analysis, we 
scale the distance as a fraction of the group's root mean square (rms) radius.

\subsection{Absolute distance to nearest neighbor}
% Including redshift

We can also calculate the distance between two galaxies by ignoring the effects 
of their relative velocities.  Instead of projecting both galaxies onto a sphere 
with radius $\bar{r}$, we instead project each onto its own sphere of radius 
\begin{equation}
    r = \frac{cz}{H_0}
\end{equation}
(This approximation for the distance is valid because all our target galaxies 
are within a redshift $z \leq 0.03$.)  We then convert to Cartesian coordinates 
and use the three-dimensional Pythagorean theorem to calculate the distance 
between the two galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Data]{SDSS Data and galaxy selection}

% SDSS
%The Sloan Digital Sky Survey Data Release 7 \citep[SDSS DR7;][]{Abazajian09} is 
%a wide-field multiband imaging and spectroscopic survey which maps approximately 
%one-quarter of the northern sky using a drift scanning technique.  Photometric 
%data in the five-band SDSS system --- $u$, $g$, $r$, $i$, and $z$ --- is taken 
%on a dedicated 2.5m telescope at the Apache Point Observatory in New Mexico 
%\citep{Fukugita96,Gunn98}.  Galaxies with Petrosian $r$-band magnitudes 
%$m_r < 17.77$ are selected for follow-up spectroscopic analysis \citep{Lupton01,
%Strauss02}.  The galaxy colors are taken from the Korean Institute for Advanced 
%Study Value-Added Galaxy Catalog \citep[KIAS-VAGC;][]{Choi10}, which contains 
%galaxies from the SDSS DR7 main sky survey based on the New York University 
%Value-Added Galaxy Catalog Data Release 7 \citep[NYU-VAGC;][]{Blanton05}.  Total 
%star formation rates (SFR) and total specific star formation rates (sSFR) are 
%from the MPA-JHU value-added catalog\footnote{Available at 
%\url{http://www.mpa-garching.mpg.de/SDSS/DR7/}}, calculated using the technique 
%described in \cite{Brinchmann04}.  
The gas-phase chemical abundances used are from Douglass, Vogeley, and Cen 
(2017, in prep.), which are calculated using the Direct $T_e$ method.  The 
galaxies' large-scale environments are based on the void catalog compiled by 
\cite{Pan12}, which is constructed from SDSS DR7 using the VoidFinder algorithm 
of \cite{Hoyle02}.  Galaxies that are located within identified voids are 
labeled as void galaxies, and those that do not fall within a void are 
designated as a wall galaxy.  The large-scale environment within 5 \hMpc of the 
edge of the main SDSS DR7 footprint cannot be described using the VoidFinder 
algorithm because the minimum diameter of a void is 10 \hMpc.  Therefore, the 
large-scale environment for galaxies within 5 \hMpc of the edge of the main 
footprint in SDSS DR7 is designated as unknown.

% photometry from KIAS-VAGC
% sSFR from MPA-JHU (Brinchmann04)
% gas-phase chemical abundances from Douglass17c

\subsection{Various samples}

% Spectroscopic sample (those with chemical abundance estimates)
% Full SDSS dwarf galaxy sample

Two galaxy samples are used in this analysis.  The largest one contains the 
11,845 dwarf galaxies spectroscopically observed in SDSS DR7.  The second sample 
contains those star-forming dwarf galaxies from the first set that have 
gas-phase chemical abundances in Douglass et al. (2017, in prep.), totaling 2506 
dwarf galaxies.  While this subset of galaxies contains only star-forming 
galaxies and is not representative of all dwarf galaxies, we show in Section 
\ref{sec:selection_effects} that this does not affect our results.

\subsection{Group catalog}

In addition to comparing the target galaxies to their nearest neighbor, we also 
look at the relationship between the dwarf galaxies' properties and their 
proximity to the nearest galaxy group.  We use the \emph{Mr18} Berlind group 
catalog \cite{Berlind06} as our source of groups, which is built on the galaxies 
in the NYU-VAGC. The catalog identifies groups using a friends-of-friends 
algorithm \citep{Huchra82} to connect galaxies with $M_r < -18$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Results]{Distance analysis and results}

Our primary objective is to compare various physical characteristics of dwarf 
galaxies against their nearest neighbors or groups to discern how the 
small-scale environment affects their evolution.

\subsection{Parameter -- distance relations}\label{sec:Relations}

The relationships between various physical parameters and the distance to the 
nearest neighbor or group for our sample of star-forming dwarf galaxies are 
shown in Figs. \ref{fig:ur}--\ref{fig:NO}.  In each of the four plots within 
each figure, we bin the galaxies in distance-space and show the average 
parameter value in each bin, to see any overall trends in the data.  We applied 
a linear regression to the data; the output of these fits are in Table 
\ref{tab:fits}.  Each of these four plots probes this distance relationship from 
a different angle.  The top left plots in Figs. \ref{fig:ur}--\ref{fig:NO} 
relate the target dwarf galaxy's parameter to the distance to its nearest 
neighbor, as defined by the the closest galaxy on the sky in \hMpc with a 
velocity difference less than 300 km/s.  The bottom left plots relate the target 
dwarf galaxy's parameter to the distance to its nearest neighbor, as defined by 
the closest galaxy on the sky in units of the virial radius of the neighbor 
galaxy with a velocity difference less than 300 km/s.  The nearest galaxy is not 
necessarily the same for these two distance measurements, as explained in Sec. 
\ref{sec:Theory_dist}.

We repeat this same analysis on the nearest groups to the target dwarf galaxy.  
The top right plots in Figs. \ref{fig:ur}--\ref{fig:NO} relate the target dwarf 
galaxy's parameter to the distance to the center of the nearest group, as 
defined as the closest group in the sky in \hMpc with a velocity difference less 
than 300 km/s.  The bottom right plots in Figs. \ref{fig:ur}--\ref{fig:NO} 
relate the target dwarf galaxy's parameter to the distance to the nearest group 
as defined by the closest group on the sky in units of the group's rms radius 
with a velocity difference of less than 300 km/s.  Groups are very rare in the 
void environment (by nature of the void environment), so the uncertainty in the 
binned values shown in these plots is much larger than in the galaxy neighbor 
plots.

These figures reveal influences from both the large-scale and small-scale 
environments, since we are identifying which galaxies reside in void regions and 
which do not.

\subsubsection{Color}

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_ur_300}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_ur_300_group}
    \caption[$u-r$ versus distance to nearest neighbor and group]{Color ($u-r$) 
    versus distance to the nearest galaxy (on the left) and nearest group (on 
    the right).  The top panel shows the color as a function of the sky 
    separation in \hMpc between the target dwarf galaxy and the neighbor, while 
    the bottom panel shows the color as a function of the closest virial 
    neighbor.  Void galaxies are shown in red, while wall galaxies are shown in 
    black and unknown in green.  We have also included the average color for the 
    galaxies after binning by distance, to discern any finer behavior in the 
    relationships.  Linear fits to the void and wall galaxies are shown in 
    orange and blue, respectively.  It is clear that the void dwarf galaxies are 
    bluer than the wall dwarf galaxies.  The nearest galaxies appear to only 
    have some affect on the dwarf galaxy's color at separations less than 0.05 
    \hMpc, or 0.1$r_{vir}$.  The closest group appears to have some affect on 
    the target dwarf galaxy's color at distances less than 0.1 \hMpc to the 
    group's center; there appears to be no relationship between a dwarf galaxy's 
    color and its distance to the nearest group as represented by the fraction 
    of the group's radius.}
    \label{fig:ur}
\end{figure}

Because of the known morphology-density relation, we expect to find that a dwarf 
galaxy's color becomes bluer as the distance to the nearest group increases, but 
we expect the dwarf galaxy's color to become bluer as the distance to the 
neighbor decreases.  Fig. \ref{fig:ur} shows very little relationship between 
the distance and color, except in the smallest distance bin.  The linear fits 
quantify this observation --- the slopes are on the order of 10$^{-3}$.  
However, within a distance of 0.05 \hMpc or 0.1$r_{vir}$, dwarf galaxies tend 
to be bluer than at further distances from their nearest neighbor.  At distances 
less than 0.1 \hMpc from the center of the nearest groups, the dwarf galaxies 
are redder than average.  However, there does not appear to be any relationship 
between a dwarf galaxy's color and its distance to the center of the nearest 
group in units of the group's rms radius.  It has been well-established that 
void galaxies tend to be bluer than galaxies in denser environments 
\citep{Grogin99,Rojas04,Patiri06,vonBendaBeckmann08,Hoyle12}; this shift is 
apparent in Fig. \ref{fig:ur}, where the void dwarf galaxies are slightly bluer 
than the wall dwarf galaxies.

\subsubsection{sSFR}

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_sSFR_300}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_sSFR_300_group}
    \caption[sSFR versus distance to nearest neighbor and group]{sSFR versus 
    distance to the nearest galaxy (on the left) and nearest group (on the 
    right).  The top panel shows the sSFR as a function of the sky separation in 
    \hMpc between the target dwarf galaxy and the neighbor, while the bottom 
    panel shows the sSFR as a function of the closest virial neighbor.  Void 
    galaxies are shown in red, while wall galaxies are shown in black and 
    unknown in green.  We have also included the average sSFR for the galaxies 
    after binning by distance, to discern any finer behavior in the 
    relationships.  Linear fits to the void and wall galaxies are shown in 
    orange and blue, respectively.  It is clear that the void dwarf galaxies 
    have higher sSFRs than the wall dwarf galaxies.  Only the neighbor galaxies 
    at separations less than 0.05 \hMpc or 0.1$r_{vir}$ appear to have some 
    affect on the dwarf galaxies' sSFR.  There appears to be no relationship 
    between a dwarf galaxy's sSFR and its distance to the nearest group.}
    \label{fig:sSFR}
\end{figure}

Following our prediction for the color-distance relations, we expect the 
sSFR to decrease with distance from the nearest neighbor and increase with 
distance from the nearest group.  We also see very little relationship between 
the distance and sSFR in Fig. \ref{fig:sSFR}, except in the smallest distance 
bin.  The linear fits quantify this observation --- the slopes are on the order 
of 10$^{-3}$.  Within a distance of 0.05 \hMpc or 0.1$r_{vir}$, dwarf galaxies 
tend to have higher sSFRs than at further distances from their nearest neighbor.  
There does not appear to be any relationship between a dwarf galaxy's sSFR and 
its distance to the center of the nearest group in either distance metric.  In 
these distance comparisons, there is a shift towards higher sSFRs in the void 
dwarf galaxies when compared to the wall dwarf galaxies, as has been observed 
before \citep{Rojas05,vonBendaBeckmann08,Moorman15,Beygu16}.

\subsubsection{Metallicity}

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_OH_300}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_OH_300_group}
    \caption[Metallicity versus distance to nearest neighbor and group]
    {Metallicity versus distance to the nearest galaxy (on the left) and nearest 
    group (on the right).  The top panel shows the metallicity as a function of 
    the sky separation in \hMpc between the target dwarf galaxy and the 
    neighbor, while the bottom panel shows the metallicity as a function of the 
    closest virial neighbor.  Void galaxies are shown in red, while wall 
    galaxies are shown in black and unknown in green.  We have also included the 
    average metallicity for the galaxies after binning by distance, to discern 
    any finer behavior in the relationships.  Linear fits to the void and wall 
    galaxies are shown in orange and blue, respectively.  It is clear that the 
    void dwarf galaxies have higher metallicities than the wall dwarf galaxies.  
    Only the neighbor galaxies at separations less than 0.05 \hMpc or 
    0.1$r_{vir}$ appear to have some affect on the dwarf galaxies' metallicity.  
    When measuring the distance to the nearest group in \hMpc, it appears that 
    the dwarf galaxies have lower metallicities at distances less than 0.1 \hMpc 
    from their nearest group.}
    \label{fig:OH}
\end{figure}

Based on our hypothesis that galaxies would be bluer and have higher sSFRs at 
small distances to their nearest neighbors, we anticipated that the metallicity 
of the galaxies would decrease with increasing distance.  As before, we only see 
a relationship between the distance and metallicity in the smallest distance 
bin in Fig. \ref{fig:OH}.  The linear fits quantify this observation --- the 
slopes are on the order of 10$^{-2}$.  Within a distance of 0.05 \hMpc or 
0.1$r_{vir}$, dwarf galaxies tend to have higher metallicities than at further 
distances from their nearest neighbor.  At distances less than 0.1 \hMpc from 
the center of the nearest group, dwarf galaxies might have lower than average 
metallicities.  However, this could be an erroneous conclusion due to 
small-number statistics.  In these distance comparisons, there is a shift 
towards higher metallicities in the void dwarf galaxies when compared to the 
wall dwarf galaxies, as has been observed by \cite{Douglass17b} and Douglass et 
al. (2017, in prep).

\subsubsection{Nitrogen abundance}

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_NH_300}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_NH_300_group}
    \caption[N/H versus distance to nearest neighbor and group]{Gas-phase 
    nitrogen abundance versus distance to the nearest galaxy (on the left) and 
    nearest group (on the right).  The top panel shows N/H as a function of the 
    sky separation in \hMpc between the target dwarf galaxy and the neighbor, 
    while the bottom panel shows N/H as a function of the closest virial 
    neighbor.  Void galaxies are shown in red, while wall galaxies are shown in 
    black and unknown in green.  We have also included the average nitrogen 
    abundance for the galaxies after binning by distance, to discern any finer 
    behavior in the relationships.  Linear fits to the void and wall galaxies 
    are shown in orange and blue, respectively.  There is very little difference 
    in N/H between the two large-scale environments.  Only the neighbor galaxies 
    at separations less than 0.1$r_{vir}$ appear to have some affect on the 
    dwarf galaxies' nitrogen abundance.  There does not appear to be any 
    relationship between the dwarf galaxies' nitrogen abundance and the distance 
    to their nearest group.}
    \label{fig:NH}
\end{figure}

It was expected that the gas-phase nitrogen abundance would follow the same 
trend as the metallicity, to decrease with increasing distance.  As observed 
with the other parameters, Fig. \ref{fig:NH} shows very little relationship 
between the distance and N/H, except in the smallest distance bin for the 
closest virial neighbor.  The linear fits quantify this observation --- the 
slopes are on the order of 10$^{-2}$.  Within a distance of 0.1$r_{vir}$, dwarf 
galaxies tend to have higher nitrogen abundances than at distances further from 
their nearest neighbor.  There does not seem to be any relationship between the 
nitrogen abundance of the star-forming dwarf galaxies and distance to the center 
of the nearest group.  Unlike the shifts seen with the other parameters and what 
is observed in \cite{Douglass17b} and Douglass et al. (2017, in prep), we see no 
significant difference in the nitrogen abundance resulting from the large-scale 
environment.

\subsubsection{N/O ratio}

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_NO_300}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_NO_300_group}
    \caption[N/O versus distance to nearest neighbor and group]{N/O ratio versus 
    distance to the nearest galaxy (on the left) and nearest group (on the 
    right).  The top panel shows the N/O ratio as a function of the sky 
    separation in \hMpc between the target dwarf galaxy and the neighbor, while 
    the bottom panel shows N/O as a function of the closest virial neighbor.  
    Void galaxies are shown in red, while wall galaxies are shown in black and 
    unknown in green.  We have also included the average nitrogen abundance for 
    the galaxies after binning by distance, to discern any finer behavior in the 
    relationships.  Linear fits to the void and wall galaxies are shown in 
    orange and blue, respectively.  The void dwarf galaxies have lower N/O 
    ratios than the wall dwarf galaxies, but there is no distinct relationship 
    between the distance to the nearest neighbor and the N/O ratio.  The N/O 
    ratio might be higher in dwarf galaxies within 0.05 \hMpc of the center of 
    the closest group.}
    \label{fig:NO}
\end{figure}

We do not expect any influence on the relative synthesis of oxygen and nitrogen 
from the proximity to a nearest neighbor since nucleosynthesis is a physical 
process unaffected by external influences.  Unlike the other parameters studied, 
Fig. \ref{fig:NO} shows that the N/O ratio does not have any relationship with 
the distance to a nearest neighbor at any separation.  This is reflected in the 
linear fits to the data --- the slopes are on the order of 10$^{-2}$ and 
smaller.  When looking at the relationship between the N/O ratio and the 
distance to the nearest group, though, the N/O ratio might be higher than 
average in galaxies within 0.1 \hMpc of the group's center.  The shift towards 
lower N/O ratios in star-forming void dwarf galaxies is readily apparent, as 
found by \cite{Douglass17b} and Douglass et al. (2017, in prep).

\subsection{Linear fit parameters}

\begin{table}
    \begin{tabular}{lcccc}
        Property & Slope (wall) & Slope (void) & Intercept (wall) & Intercept (void)\\
        \hline
        \hline
        \multicolumn{5}{c}{Nearest galaxy by distance}\\
        \hline
        $u-r$ & $1.25\pm 1.17\times 10^{-3}$ & $-9.14\pm 0.56\times 10^{-3}$ & $1.15\pm 0.00076$ & $1.09\pm0.00049$\\
        sSFR & $2.07\pm 0.18\times 10^{-2}$ & $-0.16\pm 0.08\times 10^{-3}$ & $-9.45\pm 0.001$ & $-9.34\pm 0.0007$\\
        \OH & $5.35\pm 0.17\times 10^{-2}$ & $0.33\pm 0.08\times 10^{-2}$ & $7.75\pm 0.0011$ & $7.86\pm 0.0007$\\
        \NH & $1.64\pm 0.11\times 10^{-2}$ & $0.62\pm 0.06\times 10^{-2}$ & $6.53\pm 0.0007$ & $6.52\pm 0.0006$\\
        \NO & $-3.70\pm 0.13\times 10^{-2}$ & $0.29\pm 0.07\times 10^{-2}$ & $-1.23\pm 0.0009$ & $-1.35\pm 0.0006$\\
        \hline
        \multicolumn{5}{c}{Nearest galaxy by fraction of virial radius}\\
        \hline
        $u-r$ & $-3.81\pm 0.04\times 10^{-3}$ & $-0.46\pm 0.01\times 10^{-3}$ & $1.16\pm 0.00014$ & $1.09\pm0.00008$\\
        sSFR & $5.96\pm 0.06\times 10^{-3}$ & $-0.41\pm 0.02\times 10^{-3}$ & $-9.45\pm 0.0002$ & $-9.33\pm 0.0001$\\
        \OH & $1.41\pm 0.005\times 10^{-2}$ & $0.12\pm 0.08\times 10^{-2}$ & $7.74\pm 0.0002$ & $7.86\pm 0.0001$\\
        \NH & $1.82\pm 0.11\times 10^{-3}$ & $-1.14\pm 0.002\times 10^{-3}$ & $6.53\pm 0.0001$ & $6.52\pm 0.0001$\\
        \NO & $-1.22\pm 0.004\times 10^{-2}$ & $-0.41\pm 0.02\times 10^{-3}$ & $-1.21\pm 0.0002$ & $-1.33\pm 0.0001$\\
        \hline
        \multicolumn{5}{c}{Nearest group by distance}\\
        \hline
        $u-r$ & $-2.53\pm 0.01\times 10^{-3}$ & $1.34\pm 0.01\times 10^{-3}$ & $1.16\pm 0.00008$ & $1.08\pm0.0001$\\
        sSFR & $7.77\pm 0.02\times 10^{-3}$ & $-1.68\pm 0.02\times 10^{-3}$ & $-9.48\pm 0.001$ & $-9.33\pm 0.0001$\\
        \OH & $3.65\pm 0.16\times 10^{-4}$ & $2.06\pm 0.19\times 10^{-4}$ & $7.77\pm 0.0001$ & $7.86\pm 0.0001$\\
        \NH & $4.84\pm 0.01\times 10^{-3}$ & $6.30\pm 0.01\times 10^{-3}$ & $6.51\pm 0.00008$ & $6.49\pm 0.0001$\\
        \NO & $4.47\pm 0.01\times 10^{-3}$ & $6.09\pm 0.01\times 10^{-3}$ & $-1.26\pm 0.0001$ & $-1.38\pm 0.0001$\\
        \hline
        \multicolumn{5}{c}{Nearest group by fraction of group radius}\\
        \hline
        $u-r$ & $-1.00\pm 0.002\times 10^{-3}$ & $0.90\pm 0.002\times 10^{-3}$ & $1.16\pm 0.00003$ & $1.07\pm0.00003$\\
        sSFR & $2.70\pm 0.002\times 10^{-3}$ & $-1.71\pm 0.002\times 10^{-3}$ & $-9.48\pm 0.00005$ & $-9.31\pm 0.00005$\\
        \OH & $8.81\pm 0.02\times 10^{-4}$ & $-8.74\pm 0.19\times 10^{-4}$ & $7.76\pm 0.00005$ & $7.88\pm 0.00005$\\
        \NH & $2.31\pm 0.001\times 10^{-3}$ & $2.47\pm 0.001\times 10^{-3}$ & $6.49\pm 0.00003$ & $6.47\pm 0.00004$\\
        \NO & $1.43\pm 0.002\times 10^{-3}$ & $3.35\pm 0.002\times 10^{-3}$ & $-1.26\pm 0.00005$ & $-1.40\pm 0.00005$
    \end{tabular}
    \caption[Fit parameters of properties versus distances]{Linear fit 
    parameters to various properties of the target dwarf galaxies by their 
    distances to the nearest galaxy in units of \hMpc, the nearest galaxy in 
    units of the neighbor's virial radius, the center of the nearest group in 
    units of \hMpc, and the nearest group in units of the group's rms radius; 
    all objects must be within 300 km/s of the target galaxy.  The slopes are 
    all negligible, quantifying the observations made that the proximity to a 
    galaxy or group has little influence on a dwarf galaxy's evolution.}
    \label{tab:fits}
\end{table}

To quantify the results shown in Figs. \ref{fig:ur}--\ref{fig:NO}, we calculate 
the parameters for the best linear fit.  Any slope of significant magnitude 
shows an overall correlation between a given physical parameter and the galaxy's 
distance to its nearest neighbor or group.  The results of this analysis are 
listed in Table \ref{tab:fits}.  These slopes reflect the observations described 
in Section \ref{sec:Relations}: there is no correlation between the distance to 
the nearest neighbor and a galaxy's color, sSFR, or gas-phase chemical 
abundances.  This analysis does not capture any variations within the range of 
distances for all dwarf galaxies included in this study.

\subsection{Selection effects}\label{sec:selection_effects}

% Peculiar velocity
\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_ur_150}
    \includegraphics[width=0.49\textwidth]{Images/smallScaleEnvironment/dwarf_ur_600}
    \caption[Sensitivity to peculiar velocity maximum]{Color versus distance to 
    nearest galaxy (in units of \hMpc on top and virial radii on bottom), with a 
    maximum allowed peculiar velocity of 150 km/s in the left panel and 600 km/s 
    in the right panel.  Void galaxies are shown in red, wall in black, and 
    unknown in green.  The galaxies are binned by distance to tease out any 
    trends at smaller distance scales; linear fits to the void and wall galaxies 
    are shown in orange and blue, respectively.  When compared with the two 
    plots in the left panel of Fig. \ref{fig:ur}, we see that there is no 
    significant influence on our results from the choice of maximum peculiar 
    velocity allowed.}
    \label{fig:ur_vpeculiar}
\end{figure}

We test two components of our nearest neighbor criteria to understand how 
sensitive our results are to any initial conditions.  The first parameter we 
discuss is the sensitivity of our results on the maximum peculiar velocity to 
define a match.  Throughout our analysis, we use 300 km/s as the maximum 
velocity separation allowed between the target galaxy and its nearest neighbor 
or group.  We look at how this affects our results by repeating the analysis 
with maximum velocities of 150 km/s and 600 km/s \citep[the criteria used in][]
{Hwang10,Guo11}.  The results of this comparison on the color of the galaxies 
can be seen in Fig. \ref{fig:ur_vpeculiar}.  The nearest neighbors in the 
left-hand panel are restricted to a maximum peculiar velocity of 150 km/s, while 
those in the right panel are restricted to 600 km/s.  When compared to each 
other and to the left-hand panel of Fig. \ref{fig:ur}, it is clear that our 
choice of maximum peculiar velocity has no affect on the results of the study.

% Spectroscopic sample v. complete sample
\begin{figure}
    \includegraphics[width=0.5\textwidth]{Images/smallScaleEnvironment/ALLdwarf_ur_300}
    \caption[Color versus distance of full SDSS dwarf population]{Color versus 
    distance to the nearest galaxy (in units of \hMpc on top and virial radii on 
    the bottom) for the entire dwarf galaxy population in SDSS DR7.  When 
    compared to the left-hand panel of Fig. \ref{fig:ur}, we see that there is 
    no difference in the results of the analysis by studying only star-forming 
    galaxies with sufficient detection of the various emission lines necessary 
    to estimate the gas-phase chemical abundances.}
    \label{fig:ur_allDwarf}
\end{figure}

We also test the sensitivity of our results to the population of galaxies being 
studied.  Because we want to look at the relationship between distance and the 
gas-phase chemical abundances of the dwarf galaxies, our sample is limited to 
star-forming dwarf galaxies with detected emission lines necessary for 
estimation of the chemical abundances with the Direct $T_e$ method 
\citep[see][for more details]{Douglass17a}.  We perform the same distance 
analysis on all dwarf galaxies detected in SDSS DR7 with respect to their 
color, to understand how our results depend on our sample.  When we compare Fig. 
\ref{fig:ur_allDwarf} with the left-hand panel of Fig. \ref{fig:ur}, we see that 
there is no difference in the correlation between color and distance to the 
nearest neighbor.  It is clear that our selection bias to star-forming dwarf 
galaxies does not influence any trends we observe in our analysis.

% Absolute magnitude distribution of nearest neighbors
\begin{figure}
    \includegraphics[width=\textwidth]{Images/smallScaleEnvironment/Mr_distribution}
    \caption[Distribution of absolute magnitudes of nearest neighbors]{The 
    distributions in absolute magnitude (left panel) and absolute magnitude 
    relative to the target galaxy (right panel) of the nearest neighbor 
    galaxies.  The top row includes those nearest neighbors in units of \hMpc, 
    and the bottom row consists of the nearest neighbors in units of the 
    neighbor's virial radius.  Comparing the two plots in the right panel, we 
    see that the closest galaxy to the target dwarf galaxies is often of equal 
    or fainter magnitude than the target galaxy.  Alternatively, using the 
    virial radius of the neighbor galaxy as a measure of distance often finds a 
    brighter galaxy than the target galaxy, as the bottom right plot shows.}
    \label{fig:Mr_dist}
\end{figure}

For about 41\% of our dwarf galaxy sample, the two different metrics by which to 
define the nearest neighbor (minimum in units of \hMpc or virial radius of the 
nearest neighbor) return different neighboring galaxies.  Fig. \ref{fig:Mr_dist} 
compares the absolute magnitude distributions of these two nearest neighbor 
populations.  The left panel shows the distribution of the absolute magnitudes 
of the nearest neighbor galaxies, while the right panel shows the distribution 
in absolute magnitude of the nearest neighbor galaxy relative to its target 
galaxy.  The top row includes those nearest neighbors in units of \hMpc, and the 
bottom row consists of the nearest neighbors in units of the neighbor's virial 
radius.  Comparing the two plots in the right panel, we see that the closest 
galaxy to the target dwarf galaxy is often of equal or fainter magnitude than 
the target galaxy.  Alternatively, using the virial radius of the neighbor 
galaxy as a measure of the distance often finds a brighter galaxy than the 
target galaxy, as the bottom right plot in Fig. \ref{fig:Mr_dist} shows.  By 
using these two different distance metrics, we are able to probe the 
relationship between a dwarf galaxy and its nearest neighbor and nearest dark 
matter halo.

\subsection{Including redshift in the distance}
% Evidence of mergers

\begin{figure}
    \includegraphics[width=0.5\textwidth]{Images/smallScaleEnvironment/dwarf_ur_xyz}
    \caption[Color versus distance calculated with redshift]{Color versus 
    distance to the nearest galaxy (in units of \hMpc on top and virial radii on 
    the bottom) for the star-forming dwarf galaxies.  The redshift is included 
    when calculating the distance to the nearest neighbor.  While there is still 
    no correlation between distance and color, we do note that there is a gap in 
    the distribution of galaxies around a distance of 0.05 \hMpc or 0.1$r_{vir}$ 
    from the nearest neighbor.}
    \label{fig:ur_xyz}
\end{figure}

Realizing that the peculiar velocity, which is included in a galaxy's redshift, 
can be significant for galaxies in groups, we have been careful to avoid 
calculating the distance between objects with the redshift.  However, we are 
interested to see how the inclusion of redshift in the distance calculations 
affects the results of our analysis.  Therefore, we repeat the same analysis on 
the relationship between color and distance, but this time we include redshift 
as a third component in the distance calculations.  Consequently, we no longer 
limit our sample by a peculiar velocity separation.  Fig. \ref{fig:ur_xyz} shows 
the relationship between these distances and the color of the star-forming dwarf 
galaxies.  When compared to Fig. \ref{fig:ur}, we see that there is no change in 
the correlation between distance and color for the galaxies.

We note the existence of a gap in the distribution of galaxies around a distance 
of 0.05 \hMpc or 0.1$r_{vir}$ from the nearest neighbor that is not present in 
Fig. \ref{fig:ur}.  When we incorporate the redshift into the distance 
calculations, the resulting distance between galaxies which have small sky 
separations but larger redshift separations is much larger than those with 
larger sky separations and little redshift separation.  As a result, galaxies 
which were originally close to the color axis in Fig. \ref{fig:ur} will be 
moved to much larger distances, while the locations of those which were 
originally further from the color axis will change much less.  The dwarf 
galaxies which remain close to the color axis in Fig. \ref{fig:ur_xyz} must, 
therefore, have small sky separation and almost no difference in their peculiar 
velocities.  We surmise that these represent merging systems, which future 
visual inspection will help to confirm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Discussion]{Discussion}

We see no relationship between a dwarf galaxy's color, sSFR, or gas-phase 
chemical abundances and its distance to the nearest galaxy or group beyond the 
target galaxy's immediate vicinity, implying that the small-scale environment 
($\sim$1 \hMpc) does not significantly influence a dwarf galaxy's evolution.  
This is in contrast to the large-scale environment, which we have seen to 
influence the formation and evolution of dwarf galaxies.

Only those galaxies within 0.05 \hMpc and 0.1$r_{vir}$ of a nearest neighbor, or 
those within 0.1 \hMpc of the nearest group appear to deviate from the average 
galaxy values.  The target galaxies within this proximity of their nearest 
neighbor tend to be bluer, have a higher sSFR, and have higher oxygen and 
nitrogen abundances.  Based on the shift in the distribution of galaxies seen in 
Fig. \ref{fig:ur_xyz}, these galaxies might be merging or strongly interacting 
with their nearest neighbor.  If so, this provides evidence that galaxy 
interactions result in a burst of star formation that increases the gas-phase 
chemical abundances of the dwarf galaxies.  Because merging galaxies share the 
same dark matter halo, it appears that the sharing of a dark matter halo has 
more influence on the evolution of a galaxy than the distance to its nearest 
neighbor.

In contrast, dwarf galaxies within 0.1 \hMpc of the center of the nearest group 
are redder, have lower oxygen abundances (O/H), and have higher N/O ratios than 
average.  Being so close to the center of a group seems to prevent more recent 
episodes of star formation.  Due to their proximity to the group center, it is 
also likely that these dwarf galaxies are not able to retain as much of their 
heavy elements as a more isolated galaxy, thereby reducing their gas-phase 
oxygen abundance (and increasing their N/O ratio).

\subsection{Comparison to previous results}

The influence on the gas-phase oxygen abundance within 0.05 \hMpc and 
0.1$r_{vir}$ agrees with the results of \cite{Shields91,Pustilnik06,Cooper08,
Ellison09,Pustilnik11a,Pustilnik14}, and \cite{SanchezAlmeida16}, which all find 
that galaxies with higher metallicities preferentially reside in denser regions.  
Work by \cite{Rupke08} finds that interacting galaxies have suppressed 
metallicities due to interaction- or merger-induced gas flows into the galaxy 
centers.

A study combining the effects of interactions and the large-scale environment is 
presented in \cite{Park09}.  They find two characteristic distances within which 
the behavior of the target galaxy changes: 0.05$r_{vir}$ and $r_{vir}$ of the 
neighboring galaxy.  Our results seem to confirm the significance of distances 
out to 0.05$r_{vir}$, while we see no significant change around the virial 
radius of the neighboring galaxy.  While they only look at galaxies with 
$M_r < -19$ and limit the neighbors to be at least half a magnitude brighter 
than the target, \cite{Park09} find that the morphology and luminosity play a 
significant role in these relationships.  Of particular interest is their 
observation that star formation increases in late type galaxies when their 
nearest neighbor is also of late type.  This is the same behavior we see in our 
sample of dwarf galaxies at distances less than 0.1$r_{vir}$.  With all our 
target galaxies actively forming stars, and more than half of their nearest 
neighbors also dwarf galaxies, it is most likely that our galaxy pairs are also 
of the late-late type.  Based on the results of \cite{Park09}, the deviations we 
see for those galaxies with nearest neighbors within 0.05 \hMpc and 
0.1$r_{vir}$ warrant further study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Conclusions]{Conclusions}

Using the star-forming dwarf galaxies in the SDSS DR7 sample with gas-phase 
chemical abundances from Douglass et al. (2017, in prep), we investigate the 
influence of the small-scale environment on the evolution of dwarf galaxies.  
From the $\sim$2000 galaxies in the sample, there only appears to be an effect 
from a neighboring galaxy within 0.05 \hMpc or 0.1$r_{vir}$.  The proximity of a 
group seems to only affect the target dwarf galaxy if it is within 0.1 \hMpc.  
Thus, the small-scale ($\sim 1$ \hMpc) environment does not appear to strongly 
influence the evolution of dwarf galaxies.

We examine the relationship between distance to the nearest neighbor or group 
and the target galaxy's color, sSFR, and gas-phase chemical abundances.  We find 
that, for those galaxies with a neighbor within 0.05 \hMpc or 0.1$r_{vir}$, the 
dwarf galaxies are bluer, have a higher sSFR, and have higher oxygen and 
nitrogen abundances than average.  In contrast, dwarf galaxies within 0.1 \hMpc 
of the center of the closest group are redder, have lower oxygen abundances, and 
have higher N/O ratios than average.  These results do not depend on the maximum 
peculiar velocity difference or on the sample (star-forming versus all 
galaxies).

When we incorporate the redshift into the distance calculations, we find that 
those galaxies within 0.1$r_{vir}$ are most likely mergers or strongly 
interacting with their nearest neighbor.  This matches the results of 
\cite{Park09}, who find that late-late type galaxy pairs within 0.05$r_{vir}$ 
are bluer and have higher star formation rates.  These merging galaxies likely 
share the same dark matter halo, indicating that the dark matter halo is more 
influential on a galaxy's evolution than its distance to the nearest neighbor.

Further analysis of this study should include comparing the properties of the 
target galaxies with the nearest neighbors' properties as a function of 
distance \citep[``galactic conformity'';][]{Weinmann06}.  The gas-phase chemical 
abundances between a galaxy and its nearest neighbor could be strongly 
correlated.