### 2017-06-16 KellyDouglass/Thesis/Tables/Paper3/results_table_short.tex
Updated Paper 3

\begin{sidewaystable}

    \begin{tabular}{cccccccccccc}
    Index\footnote{KIAS-VAGC galaxy index number} & R.A. & Decl. & Redshift & $M_r$ & \multicolumn{2}{c}{$12 + \log \left( \frac{\text{O}}{\text{H}} \right)$} & \multicolumn{2}{c}{$12 + \log \left( \frac{\text{N}}{\text{H}} \right)$} & \multicolumn{2}{c}{$\log \left( \frac{\text{N}}{\text{O}} \right)$} & Void/Wall\\
    \hline
	40726 & \RA{11}{25}{52}{.10} & -\dec{00}{39}{41}{.76} & 0.0187 & -16.73 & 8.18 & $\pm$0.07 & 6.70 & $\pm$0.07 & -1.49 & $\pm$0.10 & Wall \\
	40916 & \RA{11}{52}{47}{.52} & -\dec{00}{40}{07}{.64} & 0.0046 & -13.36 & 8.09 & $\pm$0.02 & 5.90 & $\pm$0.02 & -2.18 & $\pm$0.03 & Wall \\
	41257 & \RA{12}{41}{12}{.41} & -\dec{00}{45}{24}{.55} & 0.0113 & -16.87 & 8.13 & $\pm$0.20 & 6.98 & $\pm$0.18 & -1.15 & $\pm$0.27 & Wall \\
	42296 & \RA{14}{39}{50}{.03} & -\dec{00}{42}{22}{.85} & 0.0060 & -13.77 & 7.99 & $\pm$0.10 & 6.67 & $\pm$0.10 & -1.32 & $\pm$0.14 & Wall \\
	42829 & \RA{15}{40}{18}{.50} & -\dec{00}{48}{45}{.04} & 0.0125 & -16.71 & 8.06 & $\pm$0.31 & 6.95 & $\pm$0.28 & -1.11 & $\pm$0.42 & Wall
    \end{tabular}

    \caption[Analyzed dwarf galaxies]{Five of the 1920 dwarf galaxies analyzed from SDSS DR7.  The flux values for all required emission lines can be found in the MPA-JHU value-added catalog.  Abundance values are calculated using the direct $T_e$ method and the O$^+$ approximation defined in Sec. \ref{sec:Oplus_approx}, with error estimates via a Monte Carlo method.  The void catalog of \cite{Pan12} is used to classify the galaxies as either Void or Wall.  If a galaxy is located too close to the boundary of the SDSS to identify whether or not it is inside a void, it is labeled as Uncertain.  (This table is available in its entirety in machine-readable format.)}

\label{tab:Results_P3}

\end{sidewaystable}